---
title: "Lab 15 Team"
author: "Julia Foley"
date: "12/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(purrr)
options(na.action = na.warn)
movies <- read.csv("filmtv_movies - ENG.csv")
movies <- movies %>% filter(votes > 20) 
```


# Team 

## Question 

In which genre should Netflix produce a film in order to get the highest ratings? 

## Who Domain Expert is

The P-Hackers have been hired by Netflix to help them maximize the number of awards they receive for their original films. To do this, they want to know which genres of movies have been increasing in popularity in recent years. They hope to reveal a trend that will enable them to make intelligent production decisions that are more in line with modern moviegoers.   

## Why Questions are Important/Interesting

This is important because it can help Netflix determine what kind of film they should make. It is interesting to see if one genre has significantly higher rankings than  a different genre. Additionally, it will be interesting to compare genre to rating because the data could change which films are being produced. The question is important because it impacts what type of film is being made and it impacts the film industry in general. Additionally, it is important because if Netflix can produce films with higher rankings, more money will be made off of those films and therefore, Netflix will be more successful. Netflix doesn't want to take down movies that people love or put up movies that people hate because they will lose subscribers so, they must make sure that the films with the highest ratings are on Netflix.

## What Data We are Using

This data is a collection of information about films found at https://www.filmtv.it. The 12 columns found in the data are as follows:

```{r,echo=FALSE}
print(colnames(movies))
```

## Conclusion

## Recommendations 

## Findings


# Individual 

## Andres Acevedo

```{r, echo = FALSE, warning = FALSE}
movies <- movies %>% filter(genre %in% c('Drama', 'Comedy', 'Thriller', 'Documentary', 'Action', 'Horror', 'Adventure', 'Fantasy', 'Animation', 'Western'), year %in% c(1990:2019)) %>% arrange(year)

movie2 <- movies %>% group_by(genre, year) %>% summarize(year_avg = mean(avg_vote))

movie_dist <- movies %>% group_by(genre, year) %>% summarize(count = n())

movie_year <- movies %>% group_by(year) %>% summarize(movie_count = n()) %>% select(year, movie_count)

movie_dist <- full_join(movie_dist, movie_year) %>% mutate(prop = count/movie_count)
```

```{r, echp = FALSE. warning = FALSE}
ggplot(movie2, aes(x = year, y = year_avg, col = genre)) + geom_line() + facet_wrap(~genre) + theme_minimal() + theme(axis.text.x=element_text(angle = 45)) + ggtitle('Avg Rating by Genre Since 1990') + xlab('Year') + ylab('Average Rating')
```

```{r, echo = FALSE, warning = FALSE}
ggplot(movie_dist, aes(x = year, y = prop, col = genre)) + geom_line() + facet_wrap(~genre) + facet_wrap(~genre) + theme_minimal() + theme(axis.text.x=element_text(angle= 45)) + ggtitle('Genre as a Proportion of Total Films In a Year Since 1990') + xlab('Year') + ylab('Proportion')
```

### Subquestion 
How has the production of different genres of movies changed over time. Is there a clear alltime favorite type of movie for studios to produce? If so, what is it?

### Why Important/Interesting 
This question can give us a better understanding of how the behavior of production companies has changed over the years. Are they making movies in line with consumer desire? Does the distribution of movie genres make sense given the average rating of movies of a particular genre in a given year?

### New Tools Used
I utilized many dplyr functions as well as join functions to get all of the data I needed into a single tibble. In addition to all of this, I used linear models to fit a predictive line to the my data. 

### Conclusion/Answer
I analyzed the stability of the average rating of certain film genres. Volatility in these numbers over the years would indicate that there are ebbs and flows to the public's desire for different types of films. By looking for high performing genres who's average rating changed little over the described period, I was able to identify Documentary films, Animated Films, and Dramas as the three genres that have the least variability and the highest average ratings from 1990 to 2019. However, based on my data, there is a divergence between the performance of Documentary films and Animated films and the representation of these genres in the overall film pool. My recommendation to Netflix would be to produce more Documentary films and more Animated films. They consistently perform at a higher level than most other genres of film.  I was surprised to find that despite their relatively poor performance, comedies make up the highest percentage of films produced since 1990. 

### How Relate to Original Question
As a group, we were interested in seeing which genres of movies perform most highly with audiences. From this, another question arose naturally. If certain kinds of films perform better with audiences, are movie studios actually focusing their resources on the production of those genres? I wanted to see if there was any divergence between what people want, and what's actually getting made.

### Ethnical Implications
It's not always good to give people what they want. States regulate the sale of cigarettes because they're known to be hazardous to the health of the user. If Michael Bay directed all movies, audiences would be missing valuable points of view that other directors could provide. If we recommend that studios put profit and award nominations above all else, we may be asking them to sacrafice the opportunity to tell other stories that may not appeal as strongly to audiences but that benefit society by being made at all.

## Julia Foley

### Subquestion 

What affect does duration have on average rating? 

### Why Important/Interesting 

This question is interesting because it will be interesting to see if duration affects the average rating. Would people score a movie lower because the film was too long or does duration not affect ratings? I think duration would impact the average ratings and therefore, I am interested to see what the relationship is between average rating and duration. Additionally, it is important because it provides further depth into the overall question. Additionally, it will help the movie producer decide how long of movies they should produce. 


```{r}

movies_duration <- movies %>% group_by(genre) %>% summarise(duration = mean(duration)) %>% filter(duration > 75 & duration < 150)

movies_ratings <- movies %>% group_by(genre) %>% summarise(rating = mean(avg_vote))

movies_join <- movies_ratings %>% left_join(movies_duration, by = "genre")
```

```{r, se = FALSE}
ggplot(data = movies_join, mapping = aes(x = duration, y = rating, color = genre)) +
  geom_count() + 
  ggtitle("Average Duration v. Average Rating by Genre") 
```


```{r}
movies1 <- movies %>% filter(duration < 250)

movies_mod <- lm(avg_vote ~ duration, data = movies1)

moviesgrid <- movies1 %>% data_grid(duration)

moviesgrid <- moviesgrid %>% add_predictions(movies_mod) 

movies_mod
```


```{r}
ggplot(movies1, aes(duration)) +
  geom_point(aes(y = avg_vote)) +
  geom_line(aes(y = pred), data = moviesgrid, color = 'blue', size = 1) +
  ggtitle("Average Rating v. Duration") +
  labs(y = "Average Rating")
```

### New Tools Used

I used left_join in order to join the dataset with the average ratings and average duration by genre. Additionally, I used lm to create a linear model. Next, I used data-grid() to create a grid. Then, I added predictions using add_predictions. I used the linear model material that we learned in chapters 22-25. 

### Conclusion/Answer

There is a positive correlation between duration and average rating. Average rating increases as duration increases given by the equation Average rating = 0.0102(duration) + 5.2461. From the first graph, it appears that History has a long duration and an average rating of about 6.5. Mélo has the highest ratings of just over 7.5 even though it has an average duration of about 115 minutes. There is a small increase in rating as duration increase. It is important to note that the slope is relatively small but, since average rating is on a scale of 1-10 and duration is on a scale of 1-250, it makes sense that the slope would be small. It is hard to conclude that there is a linear relationship when each genre is averaged, however, there is a positive linear relationship between average rating and duration under 250. Overall, the data and linear model show that there is a positive relationship between average rating and duration. 

### How Relate to Original Question

This relates to the original question because it adds to the depth of knowledge about the relationship between genre and average rating. You can see that certain genres have higher durations and higher scores or lower durations and lower scores. The first graph shows that when the average ratings and duration are averaged by genre, there is no clear linear relationship between average ratings and duration. The graphs show that there is a relationship between duration and average rating which contributes to our overall question because it provides further depth to what causes average rating to increase. Additionally, the first graph provides further depth into the relationship between genre and average rating. 

### Ethical Implications 

My conclusion could affect the duration of movie that is made and the type of movie that is made. Since it is shown that shorter movies have higher ratings, producers may decide to make longer movies so that their ratings will be higher. Additionally, my data shows that if a producer decides to make longer films, they should make a fantasy or war movie since those movies have a higher average rating and a high average duration. My conclusion could affect the types and durations of movies that are being produced. This is good for the producers who want to make more money but, not as good for the writers of short films or for the consumers who really love short films. 

<p>
## Declan Franklin

### Subquestion 

What living actors have simular trajectories to the late great Christopher Lee? 


### Why Important/Interesting 

Christopher Lee is one of the most decorated actors of the 20th and 21th century. He has over 280 movie credits, the most in modern cinematic era. I want to point Netflix in the direction of actors on similar career trajectories as Christopher Lee. This would benefit Netflix by showing them the actors who have the will to keep producing movies, and who deserves the high paying lengthy contracts. 
 

### New Tools Used

To reveal these trajectories, I'll be using the stringr library to pick out the actors from the movies; as well as the modelr package to create linear models.


### Conclusion/Answer

```{r, echo=FALSE, message=FALSE, warning=FALSE}

christLee <- movies %>% 
  filter(str_detect(`actors`,"[C]hristopher Lee"), year>1951, year <2017) 

tomHolland <- movies %>% 
  filter(str_detect(`actors`, "[T]om Holland"), !str_detect(`actors`, "[T]om Hollander")) #Had to get rid of Tom Hollander 

chrisE <- movies %>% 
  filter(str_detect(`actors`, "[C]hris Evans"))

chrisPratt <- movies %>% 
  filter(str_detect(`actors`, "[C]hris Pratt"))

ryanReynolds <- movies %>% 
  filter(str_detect(`actors`, "[R]yan Reynolds"))

robertJr <- movies %>% 
  filter(str_detect(`actors`, "[R]obert Downey jr."))

josephL<- movies %>% 
  filter(str_detect(`actors`, "[J]oseph Gordon-Levitt"))



```



```{r, echo=FALSE, message=FALSE, warning=FALSE}
modelchrisL <- lm(avg_vote~year*votes, data = christLee)

modeltomH <- lm(avg_vote~year*votes, data = tomHolland)

modelchrisE <- lm(avg_vote~year*votes, data = chrisE )

modelchrisP <- lm(avg_vote~year*votes, data = chrisPratt)

modelryanR <- lm(avg_vote~year*votes, data = ryanReynolds)

modelrobertJr <- lm(avg_vote ~year*votes, data=robertJr)

modeljoshephL <-lm(avg_vote ~year*votes, data=josephL)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
chrisLgrid <- movies %>% 
  data_grid(avg_vote, .model=modelchrisL) %>% 
  add_predictions(modelchrisL)

tomHgrid <- movies %>% 
  data_grid(avg_vote, .model=modeltomH) %>% 
  add_predictions(modeltomH)


chrisEgrid <- movies %>% 
  data_grid(avg_vote, .model=modelchrisE) %>% 
  add_predictions(modelchrisE)

chrisPgrid <- movies %>% 
  data_grid(avg_vote, .model=modelchrisP) %>% 
  add_predictions(modelchrisP)

ryanRgrid <- movies %>% 
  data_grid(avg_vote, .model=ryanReynolds) %>% 
  add_predictions(modelryanR)

robertJgrid <- movies %>% 
  data_grid(avg_vote, .model=robertJr) %>% 
  add_predictions(modelrobertJr)

joshephLgrid <- movies %>% 
  data_grid(avg_vote, .model=josephL) %>% 
  add_predictions(modeljoshephL)

```




```{r, message=FALSE, warning=FASLE}
ggplot(data=movies) +
  geom_smooth(data =christLee, mapping = aes(y= chrisLgrid$avg_vote[1:37], x= year, colour = "Christopher Lee"), se=F) +
  geom_smooth(data=tomHolland, mapping=aes(x = year, y = tomHgrid$avg_vote[1:10], colour = "Tom Holland"), se=F) +
  geom_smooth(data=chrisPratt, mapping = aes(x= year, y = chrisPgrid$avg_vote[1:15], colour="Chris Pratt"), se=F) +
  geom_smooth(data = chrisE, mapping = aes(x = year, y= chrisEgrid$avg_vote[1:24], colour = "Chris Evans"), se=F) +
  geom_smooth(data= ryanReynolds, mapping=aes(x=year, y= ryanRgrid$avg_vote[1:25], colour = "Ryan Reynolds"), se=F) +
  geom_smooth(data= robertJr, mapping = aes(x=year, y = robertJgrid$avg_vote[1:41], colour = "Robert Downey Jr."), se=F) +
  geom_smooth(data = josephL, mapping = aes(x = year, joshephLgrid$avg_vote[1:20], colour = "Joseph Gordon-Levitt"), se=F) +
  
labs(colour = "Actors", y = "Average Popularity Votes", x= "Year") 
```



### How Relate to Original Question

This shows netflix some of the hot actors in the game. They can use this linear model to see that actors like Robert Downey Jr or Ryan Reynolds are producing gold. They could also see that actors like Tom Holland or Chris Pratt are on the come up, and should be monitored.  


### Ethnical Implications 

This data has key ethical implications because the employer at Netflix could be hiring based on his/her biased opinions. This data could help revel who the "next big thing" is, and pressure the employer to look past their biases and hire based on the numbers.


<p>

## Thomas Neal

### Subquestion 

How has the popularity of different movie genres changed with time? What genres of movies are most rapidly increasing in popularity? 

### Why Important/Interesting 

By finding what genre is most popular in each year Netflix can determine what types of movies they should license when considering older films. In addition, Netflix will know what type of movies are "trending" when considering projects for their own creation.

### Procedure 

The popularity of a movie is determined by the total amount of votes that the movie has received on FilmTV.it. The reason to look at total votes as "popularity" is because good or bad, the movie left enough of an impression to receive a vote. 

The total votes for each genre each year was then calculated. A linear model was then applied. This is represented as a wrap of total popularity across time for each genre.

 The most rapidly increasing genre in popularity will be seen by the genre with the steepest slope.


### Conclusion/Answer

```{r,echo = FALSE, warning = FALSE}
filmTV <- read.csv("filmtv_movies - ENG.csv")

topgenres <- filmTV %>% filter(filmTV$country=="United States", votes>=10) %>% select(genre, votes) %>% group_by(genre) %>% summarise(totalvotes= sum(votes)) %>% arrange(desc(totalvotes)) %>% head(12)


filmyear <- filmTV %>% filter(filmTV$country=="United States", year<2019, votes>=10, genre %in% c('Comedy','Drama','Fantasy',	'Thriller','Action','Horror'	,'Animation','Adventure','Western','Super-hero','Noir','Biography')) %>% select(-filmtv_id, -duration, -country, -director, -actors, -description, -notes) %>% group_by(year, genre) %>% summarise(totalvotes= sum(votes))



```



```{r,echo = FALSE, warning = FALSE}
popmod <- lm(totalvotes ~ year, data=filmyear)

grid <- filmyear %>% 
  data_grid(year, genre, .model = popmod) %>% 
  add_predictions(popmod)



ggplot(data=filmyear) + geom_point(aes(x=year,y=totalvotes, color=genre)) +stat_smooth(method=lm,aes(x=year,y=totalvotes), se=FALSE, color='black')+facet_wrap(filmyear$genre)  +  theme(axis.text = element_text(angle=45))


```

As can be seen the three genres with the highest slope are comedy, action, and drama. These slopes should act as predictors for the genre's popularity as time continues to pass. However, as can be seen, comedy is dropping off in popularity for 2017 and on. Because of this, I would recommend that Netflix focus on producing action or drama movies.

### New Tools Used

I used full_join in creating this data frame. I also used linear models to model the data. 

### How Relate to Original Question

This can be used as another point of data when considering what type of movie to make if looking to win awards. Often the popularity of a movie plays into whether or not it gets nominited. By knowing what genres are most popular Netflix can increase its chances of making an award-winning movie.

### Ethnical Implications 

If all movie producers used formulas like this to determine what movies that make the film scene would quickly become stale and unoriginal.

## Nisia Pinto
```{r, echo=FALSE,results=FALSE,warning=FALSE}
moviesv<-movies %>% select(year,genre,duration,avg_vote,votes) %>% filter(genre=='Action'|genre=='Adventure'|genre=='Animation'|genre=='Comedy'|genre=='Documentary'|genre=='Drama'|genre=='Drama'|genre=='Fantasy'|genre=='Horror'|genre=='Thriller'|genre=='Western')
movie_nvotes<- moviesv %>% group_by(year,genre) %>% summarize(total_nvotes = sum(votes)) 
movie_rvotes<-moviesv %>% group_by(year,genre) %>% summarize(total_avgrate = mean(avg_vote))
```

```{r, fig.width=10,echo=FALSE}
ggplot(data=movie_nvotes,aes(x=year,y=total_nvotes,color=genre))+
  geom_line(data=movie_nvotes %>% filter(genre=='Comedy'|genre=='Drama'),aes(x=year,y=total_nvotes))+
  geom_point() + ggtitle('Relationship between Genre and total Average votes')+theme_bw()
```


```{r,fig.width=10,echo=FALSE}
ggplot(data=movie_rvotes,aes(x=year,y=total_avgrate,color=genre))+
  geom_point()+
  geom_line(data=movie_rvotes %>% filter(genre=='Animation'|genre=='Documentary'),aes(x=year,y=total_avgrate))+
  ggtitle('Relationship between Genre and Number of Votes')+theme_bw()
```



```{r,fig.width=10,echo=FALSE}
grid <- movie_nvotes %>% filter(genre=='Comedy'|genre=='Drama') %>% 
  data_grid(year) 
movie_mod <- lm(total_nvotes~ year, data = movie_nvotes %>% filter(genre=='Comedy'|genre=='Drama'))
grid <- grid %>% 
  add_predictions(movie_mod) 
ggplot(data=movie_nvotes%>% filter(genre=='Comedy'|genre=='Drama'),aes(x=year,y=total_nvotes,color=genre))+
  geom_point()+
  geom_line(aes(y = pred), data = grid, colour = "blue", size = 1)+
  geom_line()+ggtitle('Prediction of Relationship between Genre and Number of Votes')+theme_bw()

```
```{r,fig.width=10,echo=FALSE}
grid <- movie_rvotes %>% filter(genre=='Documentary'|genre=='Animation') %>% 
  data_grid(year) 
movie_mod <- lm(total_avgrate~ year, data = movie_rvotes %>% filter(genre=='Documentary'|genre=='Animation'))
grid <- grid %>% 
  add_predictions(movie_mod) 
ggplot(data=movie_rvotes %>% filter(genre=='Documentary'|genre=='Drama'),aes(x=year,y=total_avgrate,color=genre))+
  geom_point()+
  geom_line(aes(y = pred), data = grid, colour = "Blue", size = 1)+
  geom_line()+ggtitle('Prediction of Relationship between Genre and Total Average Ratings')+theme_bw()

```


### Subquestion 

What is the relationship between the number of votes and the movie genre? Is quantity better than quality?

### Why Important/Interesting 

Nowadays, when choosing a movie people usually go online and see the rating and through ratings, one can identify if the movie that he/she will see is worth the time or not. However, in other cases when looking for old films that people did like at that time have a low score. I would like to show if the rating is really important when choosing a genre or we can just choose it by the popularity of the genre. Several popular movies have bad ratings and the number of votes represents the number of people that saw the movies. I believe this would be a helpful tool for Netflix to see what people are watching regardless of quality. Quality is important, but sometimes people just want to binge-watch and most movies have a rating.

### New Tools Used

In this lab, I mostly used a recently learned tool: predictions. I used lm and predictions to predict the relationship between the average number of rates, the total number of votes and genre. I used the modelr package.I used the piping methods we learnerd in the first months of classes.

### Conclusion/Answer

From 1990 to 2020 there is a decrease in the rating scores and an increase in the number of votes. There is a negative prediction for total average ratings and a positive prediction for the total number of votes.
The conclusion we were able to observe is that rating is not the ultimate factor. It would be good if Netflix decided to make more quality movies. The best way to decide what genre to choose is to be based on the number of votes rather than the rating number. However, the rating average for each genre is reducing as the number of years increases. It could mean that it is becoming less popular. On the other hand, the number of votes increases significantly. Consequently, from data, we can see that Netflix should invest more in the **Drama** genre if they want to keep popularity and in **Documentary** if their concerns are the average of the ratings. 

### How to Relate to Original Question

This question shows us the relationship between genre, number of votes and ratings. It helps us answer the original question because it will show us what type of movie genre Netflix should produce and buy according to popularity. It also explains if quantity is better than quality.

### Ethical Implications 

This information could affect how producers make movies. Most producers will focus only on the number of views so movies like this could be on Netflix. They will primarily focus on one type of genre. As a consequence, they would not evolve artistically. They would make lower quality movies as a consequence producer will produce lower quality movies in the genre that most people watch and distance themselves more from quality movies.


# Reflection

## Andres Acevedo 

## Julia Foley

For my six months after college goal, originally, I thought I wanted to go to law school but, now, I think I want to do data science. Additionally, my original five-year goal was that I wanted to work for the FBI and while, I do still want to work for the government, I do not think I want to work in the FBI anymore. I'm not really sure what I want to do with my life but, I know I'll figure it out and that data science will be in it. I have learned a lot in this course. I know know the basics of data science and how to work R. I learned a lot about creating and anaylzing graphs. I thought it was really cool that I had the chance to learn about data science. I am excited to continue as a data science major thanks to this course. If I were to give myself advice, I would say that it is okay if you don't know everything, you have a pretty great team that will help you succeed. Additionally, I would say that I should plan out more time to meet with my team in person to work on the lab. Additionally, I would tell myself to keep working through the problems and you'll eventually get it. Lastly, I would tell myself to keep working hard and keep finishing the labs early because it reduces the amount of stress you have on Friday afternoons after lecture. 

## Declan Franklin

## Thomas Neal

My six-month goal has now pivoted to making sure I do well in my spring semester classes. My five-year goal remains unchanged. From this class, I have learned I like all the questions that come with data science. I like having to interpret data and wonder why something is happenting. If I could give myself some advice it would be not to worry and to stick to the plan. I came in with a plan this semester and executed it well.

## Nisia Pinto

My six-month plan changed in the past semester. I am more environmentally friendly and aware. I want to work in a field related to waste management and water treatment. Five years from now I would like to have in my portfolio work experience in South Korea due to their technological advances in the recycling and water treatment area. I would like to have a small school to help kids in Angola, nothing too formal, but just a mean to spread knowledge. I learned a lot this semester and one of the main aspects was how to work in large groups and how to communicate. I learned R programming which I did not even know what it was previous to this class. I learned how to ask data questions and what type of information someone should look for so they could give an educated answer. If I  could give myself a piece of advice, it  would be to go more to OH and practice more R with the professor and also with my team.


# What Everyone Did 

## Andres Acevedo 

## Julia Foley

For my individual, I grouped by genre and then, summarized the mean of average rating and duration. I then, left joined the dataset in order to be able to compare the two averages. Next, I created a linear model and added predictions. Next, I graphed the data points and the linear model on a graph. Lastly, I wrote a conclusion, analyzed how it related to my original question, and its ethical implications. For the team section, I made the outline. I then worked with the rest of my team to figure out what we were studying, what dataset we were using, and who are domain expert was. Then, I wrote the Why Important/Interesting section of the team part. 

## Declan Franklin

For the team part, I helped create our question which the lab is based on. I wrote on the back ground of the question and explained our answer. I also wrote cleaned up and submitted the lab.

For the individual part, I created a linear model to compare some of the current hot actors to the great Christopher Lee. I used the tidyverse, stringr, modler, and purrr libraries to do so.

## Thomas Neal

For the team part, I wrote about the data and some background. For my individual part, I  found the 12 most popular genres across time. I then grouped/ filtered to show these genres the data biased off this. I then graphed each genre's popularity with respect to time. Lastly, I created a linear model for each genre to find which ones are most popular.

## Nisia Pinto

Individual: I compared the relationship between genre and  total_avearge rating and a total number of votes and gave a  conclusion to my particular case.

Team: I helped the team brainstorm ideas for the project and helped come up with some questions. I talked about the questions for the team part we all decided to implement in our report.
