---
title: "Lab 13"
author: "P-Hackers (Andres Acevedo, Julia Foley, Declan Franklin, Thomas Neal, Nisia Pinto)"
date: "11/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
earthquake <- read.csv('earthquake_affected_annual_number.csv')
naturalgas <- read.csv('natural_gas_production_total.csv')

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
   perm_mean_diffs2 <- numeric(perms)
  
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    rand_order_y <- sample(y)
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    cor_comp <- cor(x,rand_order_y)
    perm_mean_diffs2[i] <- cor_comp 
    
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
return(perm_mean_diffs2)
}

```

# Team 

### Question

### Why Interesting/Important? 

### Describe Dataset/ Preprocessing of the Data

### Summary of Answer to Overall Question

# Individual 

### Andres Acevedo

```{r, warning = FALSE, echo = FALSE}
energy_use <- read_csv("energy_use_per_person.csv")


energy <- energy_use %>% filter(country == "United States") %>% as_tibble()


year <- c(1990:2008)
energy_values <- c(7670, 7630,	7680,	7710,	7760,	7760,	7840,	7830,	7800,	7920,	8060,	7830,	7840,	7790,	7880, 7850,	7700,	7760,	7490)

energy_USA <- tibble(year, energy_values)


forest <- read_csv("forest_coverage_percent.csv")
forest <- forest %>% filter(country == "United States")

year <- c(1990:2008)
forest_values <- c(33,	33,	33,	33.1,	33.1,	33.1,	33.1,	33.1,	33.1,	33.1,	33.1,	33.2,	33.2,	33.2,	33.2,	33.3,	33.4,	33.4,	33.6)

forest_coverage <- tibble(year, forest_values)



quantile(perm_cor(1000, energy_USA$energy_values, forest_coverage$forest_values), 0.95)

correlation_data <- tibble(data)
average_corr <- map(correlation_data, mmean)
```

##### Data Set and Why 
I chose to use 2 datasets. The first containing data on the amount of power used by each person in the United States. The second containing the percentage of the United States covered in forest in a given year. I wanted to see if I could find a statistically significant correlation between these two variables.

##### Subquestion
Does the amount of electricity we use really have a direct affect on the amount of forest left in the United States or is it merely a coincidence that these trends occured together. What does cause loss of forest land? What could be directly linked to it? 

##### Hypothesis
There Will be no statistically significant link between the energy used per person and the percent of the United States covered in forest. This could cause loss of forest indirectly. Weather conditions have a more direct impact like causing forest fires.

##### Permutation Test
```{r, warning = FALSE, echo = FALSE}
data <- perm_cor(1000, energy_USA$energy_values, forest_coverage$forest_values)
```

##### Map Function
```{r, warnng = FALSE, echo = FALSE}

correlation_data <- tibble(data)

average_corr <- map(correlation_data, mean)
average_corr
```

**What Map Function is doing**
After making a tibble out of the list of 1000 correlations, I found their mean using the map function with the mean function as an argument. This helped me to see that it's most common to get almost no correlation between these variables.

##### Permutation Test
```{r, warning = FALSE, echo = FALSE}
data <- perm_cor(1000, energy_USA$energy_values, forest_coverage$forest_values)
```


**Null and Alternative Hypothesis**

Null Hypothesis: there is no correlation between energy use per person and the percentage of the United States covered in forest.

Alternative Hypothesis: there is a correlation between energy use per person and the percentage of the United States covered in forest. 

**Test Statistic**
```{r, warning = FALSE, echo = FALSE}
corr <- cor(energy_USA$energy_values, forest_coverage$forest_values)
```
The correlation coefficient was -.2973998


**Histogram**
```{r, warning = FALE, echo = FALSE}
ggplot() + geom_histogram(mapping = aes(x = data), col = 'steelblue', fill = 'black') + geom_vline(xintercept = corr, col = 'red' ) + geom_vline(xintercept = .4026999, col = 'green') + ggtitle("Distribution of Correlation Coefficient Permutations") + xlab('Correlation Coefficient(Red = 95th Percentile, Green = Test Stat)') + ylab('Count')

```

**Percentile of Real Data**
```{r, warning = FALSE, echo = FALSE}
sum(data < -.2973998)/1000
```

**Conclusion to Test and Question**

Since 9.2% of all of the correlations I took are less than my test statistic, this shows that this isn't that unlikely of an outcome from a random test. Which is what our permutation test was. The mean correlation of all of my data was .00297816  which is very close to zero. I accept my null hypothesis that are there is no significant correlation between the power usage per person and the percent of U.S. land covered in forests. This could be for many reasons other than just the amount of power people use. But more with other conditions like dryness and high temperatures or even logging because of the increased demand for housing as the population increases and development continues.



### Julia Foley

##### Data Set and Why 

I wanted to anaylze a natural disaster so I brainstormed how natural disasters could relate to energy. I chose to use the dataset about the Natural Gas Production per year and the Number of People affected by Earthquakes per year. I thought that these were interesting because I wanted to see if Natural Gas Production contributes to Earthquakes. I thought there could be a correlation because of fracking. I think these datasets will be interesting to analyze and will produce a correlation. 

##### Subquestion

Is there a correlation between the number of people affected by Earthquakes and the amount of Natural Gas produced in the last 25 years (1983-2008)? 

##### Hypothesis

I believe that there will be a positive correlation between the number of people affected by earthquakes and the amount of natural gas produced. 

##### Map Function

```{r}
earthquake <- earthquake %>% select(country, X1983, X1984, X1985, X1986, X1987, X1988, X1989, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008) %>% gather('X1983', 'X1984', 'X1985', 'X1986', 'X1987', 'X1988', 'X1989', 'X1990', 'X1991', 'X1992', 'X1993', 'X1994', 'X1995', 'X1996', 'X1997', 'X1998', 'X1999', 'X2000', 'X2001', 'X2002', 'X2003', 'X2004', 'X2005', 'X2006', 'X2007', 'X2008', key = "year", value = "value") %>% filter(!is.na(value)) %>% rename("Number affected by Earthquake" = value) 
naturalgas <- naturalgas %>% select(country, X1983, X1984, X1985, X1986, X1987, X1988, X1989, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008) %>% gather('X1983', 'X1984', 'X1985', 'X1986', 'X1987', 'X1988', 'X1989', 'X1990', 'X1991', 'X1992', 'X1993', 'X1994', 'X1995', 'X1996', 'X1997', 'X1998', 'X1999', 'X2000', 'X2001', 'X2002', 'X2003', 'X2004', 'X2005', 'X2006', 'X2007', 'X2008', key = "year", value = "value") %>% filter(!is.na(value)) %>% rename("Natural Gas Production" = value) 




# there will be a positive correlation between the number affected by earthquakes and the natural gas production 




naturalgas <- naturalgas %>% mutate(year2 = parse_number(str_sub(naturalgas$year, 2, 5)))

earthquake <- earthquake %>% mutate(year2 = parse_number(str_sub(earthquake$year, 2, 5)))


gasEarthquake <- earthquake %>% inner_join(naturalgas, by = c("country", "year2"))

```

```{r}
gasEarthquake2 <- gasEarthquake %>% select('Number affected by Earthquake', 'Natural Gas Production')
china <- gasEarthquake %>% filter(country == "China") %>% select('Number affected by Earthquake', 'Natural Gas Production')
Indonesia <- gasEarthquake %>% filter(country == "Indonesia")  %>% select('Number affected by Earthquake', 'Natural Gas Production')
US <- gasEarthquake %>% filter(country == "United States") %>% select('Number affected by Earthquake', 'Natural Gas Production') 
Pakistan <- gasEarthquake %>% filter(country == "Pakistan")  %>% select('Number affected by Earthquake', 'Natural Gas Production')
India <- gasEarthquake %>% filter(country == "India")  %>% select('Number affected by Earthquake', 'Natural Gas Production')
```


```{r}
map(gasEarthquake2, function(x){mean(x)})
```

```{r}
map(china,function(x){mean(x)})
```

```{r}
map(Indonesia, function(x){mean(x)})
```

```{r}
map(US, function(x){mean(x)})
```

```{r}
map(Pakistan,function(x){mean(x)})
```

```{r}
map(India, function(x){mean(x)})
```


**What Map Function is doing**

My map function calculates the average number of people affected by earthquakes each year and the average amount of natural gas produced in five countries (China, Indonesia, Columbia, Pakistan, India). I chose these countries because they produce a lot of Natural Gas and are somewhat near the ring of fire. 


##### Permutation Test 

**Null and Alternative Hypothesis**

My null hypothesis is cor(A,B) > 0, my alternative hypothesis is cor(A,B) < 0

**Test Statistic**

My test statistic is cor(A, B)

```{r}
perm_cor <- function(perms = 1000, n_earth, n_gas)
{
  # perms: number of permutations
  # n_gas (int): number of gas production
  # n_earth (int): number of earthquakes 
  ###############
  
    perm_mean_diffs2 <- numeric(perms)
  
  
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    
     rand_order_y <- sample(n_gas)
     
    
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    
     perm_cor_mixed <- cor(n_earth, rand_order_y, method = c("pearson"))
    perm_mean_diffs2[i] <- perm_cor_mixed
    
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(perm_mean_diffs2)
}


```

```{r}
perm_cor(10, gasEarthquake$`Number affected by Earthquake`, gasEarthquake$`Natural Gas Production`)

```

**Histogram**

```{r}
Juliahist <- hist(perm_cor(1000, gasEarthquake$`Number affected by Earthquake`, gasEarthquake$`Natural Gas Production`), main = "Number affected by Earthquakes Compared to Natural Gas Production", xlab = "Linear Correlation")

Juliahist
```


**Percentile of Real Data**

```{r}
Juliapercent <- quantile(perm_cor(1000, gasEarthquake$`Number affected by Earthquake`, gasEarthquake$`Natural Gas Production`))

Juliapercent
```


**Conclusion from Permutation test**

The permutation test shows that there is little to no correlation between the number of people affected by natural gas and natural gas production. However, the majority of camparisons do have a slight negative correlation which is different than I originally believed. My permutation test showed that in the 50th percentile, there was a -.029 correlation between the number of people affected by earthquakes and the amount of natural gas produced. This is a very small correlation so it is too negligible to say there is a correlation. Additionally, the percentiles show that the correlation switches from negative to positive which means there can not be a consistent correlation. 

##### Conclusion to Question

My original hypothesis was incorrect. There is no correlation between the number of people affected and the amount of natural gas produced. My data shows that though some companies that produce large amounts of gas have higher than average numbers of people being affected by earthquakes (China and India), overall, there is little correlation between the two. Overall, by testing the means and correlations between the number of people affected by earthquakes and the amount of natural gas produced, we can conclude that there is no correlation. This means that my hypothesis was false. Eventhough there was no correlation between the number of people affected by earthquakes and the amount of natural gas produced, there may be a correlation between the number of earthquakes and the amount of natural gas produced however, we do not have that data within the gapminder dataset. 


### Declan Franklin

##### Subquestion

##### Hypothesis

##### Map Function

**What Map Function is doing**

##### Permutation Test 

**Null and Alternative Hypothesis**

**Test Statistic**

**Histogram**

**Percentile of Real Data**

**Conclusion from Permutation test**

##### Conclusion to Question



### Thomas Neal 

##### Data Set and Why 

I have chosen the babies per woman data set. This data set shows the average amount of babies each mother had for a given year. I choose this data set because I wanted to see the average family size was decreasing in the United States.

##### Hypothesis

The is no difference between the average amount of babies born in the 1960s and the 2000s in the United States.

##### Subquestion

What is the exact difference in the average amount of babies per mother?

##### Map Function

```{r},echo=FALSE,message=FALSE
BabiesPerWomen <- read_csv("https://raw.githubusercontent.com/juliafoley/P-Hackers/master/children_per_woman_total_fertility.csv")

BabiesPerWomenUSA60 <- BabiesPerWomen %>% filter(country=="United States") %>% select(`country`,`1960`,`1961`,`1962`,`1963`,`1964`,`1965`,`1966`,`1967`,`1968`,`1969`) %>% gather(`1960`,`1961`,`1962`,`1963`,`1964`,`1965`,`1966`,`1967`,`1968`,`1969`,key="Year",value="Value")

BabiesPerWomenUSA00 <- BabiesPerWomen %>% filter(country=="United States") %>% select(`country`,`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,`2008`,`2009`) %>% gather(`2000`,`2001`,`2002`,`2003`,`2004`,`2005`,`2006`,`2007`,`2008`,`2009`,key="Year",value="Value")

BabiesPerWomenUSA <- full_join(BabiesPerWomenUSA60,BabiesPerWomenUSA00)

BabiesMean00 <- BabiesPerWomenUSA00 %>% summarise(mean=mean(Value))
BabiesMean60 <- BabiesPerWomenUSA60 %>% summarise(mean=mean(Value))
```


```{r, eval = T,echo=FALSE,message=FALSE}
perm_mean <- function(perms = 1000, all_values, n_A)  # All values of hist and cmci
{

  perm_mean_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "all values" into disjoint 
    # groups of size "nA" and "nB" respectively
    # we will take the vector, shulf it, and call n_A of the top and call it group A, then the rest will be group B
    rand_order <- sample(all_values)
    group_A <- rand_order[1:n_A] # First go to n_A
    group_B <- rand_order[n_A:length(all_values)] # Rest go here
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    
    A_bar <- mean(group_A)
    B_bar <- mean(group_B)
    
   
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1

    diff_means <- A_bar - B_bar
    perm_mean_diffs[i] <- diff_means
  }
  
  # Step 5:
  # Return the permuted mean differences vector
return(perm_mean_diffs)
  
}
```



```{r,echo=FALSE,message=FALSE}
BabiesPerWomenUSA00

aveBabies00<-BabiesPerWomenUSA00 %>% select(Value,country) %>% rename("Average Babies Per Mother in 2000s"=Value)
aveBabies60 <- BabiesPerWomenUSA60 %>% select(Value,country) %>% rename("Average Babies Per Mother in 1960s"=Value)

avebabies <- left_join(aveBabies00,aveBabies60) %>%  select(`Average Babies Per Mother in 2000s`,`Average Babies Per Mother in 1960s`)


```

```{r}


avebabies<- map(avebabies, function(x){mean(x)})
avebabies

```


```{r,echo=FALSE,message=FALSE}
Fakemean <- perm_mean(10000,BabiesPerWomenUSA$Value,10)

Realmean = BabiesMean60-BabiesMean00


ggplot() +geom_histogram(mapping = aes(x=Fakemean)) +geom_vline(xintercept = Realmean$mean,color="red")

quantile(Fakemean,Realmean$mean)

```


**What Map Function is doing**

Finding the average babies per mother in the 1960s and the average babies per mother in the 2000s.

##### Permutation Test 

**Null and Alternative Hypothesis**

My null hypothesis is mean(A)-mean(b) = 0, my alternative hypothesis is mean(A)-mean(b) ~= 0

**Test Statistic**

mean(A)-mean(b)

**Histogram**

```{r,echo=FALSE,message=FALSE}

Fakemean <- perm_mean(10000,BabiesPerWomenUSA$Value,10)

Realmean = BabiesMean60-BabiesMean00


ggplot() +geom_histogram(mapping = aes(x=Fakemean)) +geom_vline(xintercept = Realmean$mean,color="red")+labs(title = "
Distribution of permutation test data", x="Mean Difference")+theme_classic() 

quantile(Fakemean,Realmean$mean)

```

**Percentile of Real Data**

```{r,echo=FALSE,message=FALSE}


quantile(Fakemean,Realmean$mean)

```

**Conclusion from Permutation test**

With the real mean difference falling in the  99.1% percentile we reject the null Hypothesis.

##### Conclusion to Question

The amount of babies per mother is not the same between the 1960s and the 2000s. In the 1960s the average mother was having almost exactly one more child than in the 2000s.


### Nisia Pinto

```{r,echo=FALSE, warning=FALSE,results=FALSE,results=FALSE,message=FALSE}
library(tidyverse)
library(tidyr)
library(gapminder)
library(dplyr)
```

##### Data Set and Why

Environmental:planted_forest_area_ha and Energy:coal_consumption_total. I chose this data because I wanted to analyze if the world is becoming a significantly more conscious place regarding reforestation and coal consumption. I would like to know if we are decreasing carbon footprints and relying less on coal.

```{r,echo=FALSE,warning=FALSE}
coal <- read.csv('coal_consumption_total.csv')
forest<- read.csv('planted_forest_area_ha.csv')
```

```{r,echo=FALSE,warning=FALSE,results=FALSE}
coal1<- coal %>% select(country, X1965, X1966, X1967, X1968, X1969, X1970, X1971, X1972, X1973, X1974,X1975, X1976, X1977, X1978, X1979, X1980, X1981, X1982,X1983, X1984, X1985, X1986, X1987, X1988, X1989, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008) %>% gather('X1965', 'X1966', 'X1967', 'X1968', 'X1969', 'X1970', 'X1971', 'X1972', 'X1973', 'X1974','X1975', 'X1976', 'X1977', 'X1978', 'X1979', 'X1980', 'X1981', 'X1982','X1983', 'X1984', 'X1985', 'X1986', 'X1987', 'X1988', 'X1989', 'X1990', 'X1991', 'X1992', 'X1993', 'X1994', 'X1995', 'X1996', 'X1997', 'X1998', 'X1999', 'X2000', 'X2001', 'X2002', 'X2003', 'X2004', 'X2005', 'X2006', 'X2007', 'X2008', key = "year", value = "value") %>% filter(!is.na(value)) %>% rename("Total Coal Emissions" = value) 
coal1 <- coal1 %>% mutate( Year= parse_number(str_sub(coal1$year, 2, 5)))
coal1<-coal1 %>% select('country','Year','Total Coal Emissions')
coal1<-coal1 %>% filter(Year>=1990) %>% filter(!is.na(`Total Coal Emissions`))
```


```{r,echo=FALSE,warning=FALSE,results=FALSE}
forest1<- forest %>% select(country, X1990, X1991, X1992, X1993, X1994, X1995, X1996, X1997, X1998, X1999, X2000, X2001, X2002, X2003, X2004, X2005, X2006, X2007, X2008) %>% gather('X1990', 'X1991', 'X1992', 'X1993', 'X1994', 'X1995', 'X1996', 'X1997', 'X1998', 'X1999', 'X2000', 'X2001', 'X2002', 'X2003', 'X2004', 'X2005', 'X2006', 'X2007', 'X2008', key = "year", value = "value") %>% filter(!is.na(value)) %>% rename("Planted forest area (ha)" = value) 
forest1 <- forest1 %>% mutate(Year= parse_number(str_sub(forest1$year, 2, 5)))
forest1<-forest1 %>% select('country','Year','Planted forest area (ha)')
forest1<-forest1 %>% filter(Year>=1990) %>% filter(!is.na(`Planted forest area (ha)`))
```

```{r,echo=FALSE,warning=FALSE,results=FALSE,message=FALSE}
ndata<-full_join(forest1,coal1) %>% filter(!is.na(`Planted forest area (ha)`)) %>% filter(!is.na(`Total Coal Emissions`))
ndata1<-full_join(ndata,gapminder) %>% select(country,continent,`Planted forest area (ha)`,`Total Coal Emissions`) %>% filter(!is.na(`Planted forest area (ha)`)) %>% filter(!is.na(`Total Coal Emissions`))
asias<-ndata1 %>% filter(continent=='Asia')
asias<-asias %>% select(`Planted forest area (ha)`,`Total Coal Emissions`)
americas<-ndata1 %>% filter(continent=='Americas')
americas<-americas %>% select(`Planted forest area (ha)`,`Total Coal Emissions`)
africas<-ndata1 %>% filter(continent=='Africa')
africas<-africas %>% select(`Planted forest area (ha)`,`Total Coal Emissions`)
europes<-ndata1 %>% filter(continent=='Europe')
europes<-europes %>% select(`Planted forest area (ha)`,`Total Coal Emissions`)
```
##### Hypothesis

There is a negative correlation between the planted forest area and coal consumption in the world. There is an increase in planted areas and a decrease in coal consumption.

##### Subquestion

Is there a correlation between the two Planted Fores Area and Coal Consumption from 1990 to 2008?

##### Map Function
 
```{r,echo=TRUE,warning=FALSE,results=FALSE}
map(asias,function(x){sd(x)})
map(africas,function(x){sd(x)})
map(europes,function(x){sd(x)})
map(americas,function(x){sd(x)})
```
**What Map Function is doing**

This map function calculates the standard deviation of Planted forest Area and Total Coal Consumption from the mean from 1990 to 2008. This map function takes each element in the sample and returns the standard deviation.

##### Permutation Test 

**Null and Alternative Hypothesis**

Null hyphothesis is (mean(coal consumption)-mean(planted forest area))equals to zero.

Alternative hypothesis: (mean(coal consumption)-mean(planted forest area)) differen than zero.

**Test Statistic**

We were able to verify that there is a high positive correlation of 0.9327708 between the two datasets.
 

```{r,echo=FALSE,warning=FALSE,results=FALSE}

perm_cor <- function(perms = 1000, x, y)
{
   perm_mean_diffs2 <- numeric(perms)
  for (i in c(1:perms))
  {
    rand_order_y <- sample(y)
    cor_comp <- cor(x,rand_order_y)
    perm_mean_diffs2[i] <- cor_comp 
    
  }
  return(perm_mean_diffs2)
}
````

```{r,echo=FALSE}
perm<-perm_cor(1000,ndata1$`Planted forest area (ha)`,ndata1$`Total Coal Emissions`)

```

```{r,echo=FALSE,warning=FALSE,results=FALSE}
corr<-cor(ndata1$`Total Coal Emissions`,ndata1$`Planted forest area (ha)`) #0.9327708
a<-meancoal<-map(ndata1 %>% select(`Total Coal Emissions`),function(x){mean(x)})#46729447
b<-meanforest<-map(ndata1 %>% select(`Planted forest area (ha)`),function(x){mean(x)})#3753876
diff.means <-abs( mean(ndata1$`Planted forest area (ha)`) - mean(ndata1$`Total Coal Emissions`))/(46729447+3753876)
mean=diff.means/1000 #0.0008512825

```

**Histogram**

```{r,echo=FALSE,message=FALSE,warning=FALSE}
ggplot() + geom_histogram(mapping = aes(x = perm), col = 'black', fill = 'lightyellow') + geom_vline(xintercept = mean, col = 'red' )+theme_bw()+ggtitle('Permutation Test')
```

**Percentile of Data**

```{r,echo=FALSE,message=FALSE,results=TRUE}
quantile(perm)
```

**Conclusion from Permutation test**

With this permutation test, we will reject the null hypothesis, but we do not know if we should accept the alternative hypothesis. There is a difference in the means between the two datasets.

##### Conclusion to Question

There is a high correlation between the two datasets we analyzed. However, it is not negative; it is positive. Potentially, this could mean that they both decrease, or they both increase. As a consequence, my hypothesis is wrong, and both datasets are proportional.
