---
title: "Lab11_P-Hackers"
author: "Declan Franklin"
date: "11/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(babynames)
library(stringr)

popnames <- read.csv("https://query.data.world/s/whygze632j7ymm2cfy4rwcyvzvdaeq", header=TRUE, stringsAsFactors=FALSE)

as_tibble(popnames)

yob2018 <- read.csv("https://raw.githubusercontent.com/juliafoley/P-Hackers/master/yob2018.txt",header=FALSE)  %>% rename(name = 'V1', sex = 'V2', n = 'V3') %>% mutate(year=2018)


totalcount <- yob2018 %>% summarise(sum=sum(n))

yob2018 <- yob2018 %>% mutate("prop"=yob2018$n/totalcount$sum)




babynames <- babynames

totalnames <- full_join(babynames,yob2018) 


```
# Individual Plot of your Name's Popularity 

## Andres Acevedo 

```{r,echo=FALSE , warning=FALSE, message=FALSE}
variant_count <- popnames %>% 
  filter(str_detect(name, pattern = "^And")) %>% 
  group_by(year) %>% summarize(variants = n(), sum(count))

test_join <- popnames %>% group_by(year) %>% summarize(total = sum(count))

test_full_join <- variant_count %>% full_join(test_join) %>% 
mutate(prop = `sum(count)`/total)

andres <- popnames %>% filter(str_detect(name, pattern = "^A[in|n]d[re]..$")) %>% 
  group_by(year) %>% summarize(count = sum(count)) %>% 
  full_join(test_join) %>% mutate(prop = count/total)

ggplot(test_full_join) + 
  geom_line(aes(x = year, y = prop), col = "steelblue") + 
  geom_line(data = andres, aes(x = year, y = prop), col = 'red') +
  ggtitle("Time series of Firsth Three Letters (Blue) and Regex (Red) Over Time") + 
  xlab("Year") + ylab("Proportion") + theme_minimal()
```
***Regular expression***
The regular expression I used was ^A[in|n]d[re]..$, all of the common variations of my name started with either An or Ain with d always following the first letters. The next letter was always either r or e. Common variatons of my name can have additional letters at the end such as eas. 

***Name now Compared to when I was Born***
My name has decreased in popularity since I was born. You can see the exact difference in the "Your names" section of this lab. For the most part, the popularity of my name had been increasing since 1880 but began to decline around 1984. 

***How Popular is your name compared to other names that start with "And"?*** 
Based on my plot, names starting with 'And' are more popular than variations of my name 'Andres' this would make sense as there are many names starting with 'And' that are not variations of Andres. It would make sense that the use of a more strict regular expression would lead to a dataframe with fewer names and therefore a smaller proportion of total names in a given year. 

## Julia Foley

```{r,echo=FALSE , warning=FALSE, message=FALSE}
Julia1 <- totalnames %>% filter(str_detect(name, "^J[eu]+[ew]*l[iey]+a$"))
Julia2000 <- Julia1 %>% filter(year %in% 2000) %>% group_by(year) %>% summarise(prop_A = sum(prop)) # .00443
Julia2018 <- yob2018 %>% filter(str_detect(name, "^J[eu]+[ew]*l[iey]+a$"))
Juliaprop <- 2999/32032 #.09363
Julia1982 <- Julia1 %>% filter(year %in% 1982) %>% group_by(year) %>% summarise(prop_A = sum(prop)) #.0019455
Jul0018 <- 	0.00443067 - Juliaprop
Jul8200 <- 0.00194553 -	0.00443067 

Julia1a <- totalnames %>% filter(str_detect(name, "^Julia$"))
Julia2000a <- Julia1a %>% filter(year %in% 2000) %>% group_by(year) %>% summarise(prop_A = sum(prop)) # .00443
Julia2018a <- Julia1a %>% filter(year %in% 2018) %>% group_by(year) %>% summarise(prop_A = sum(prop)) 
Julia1982a <- Julia1a %>% filter(year %in% 1982) %>% group_by(year) 
Julia1982suma <- Julia1982a %>% summarise(prop_A = sum(prop)) #.0019455
Jul0018a <- 0.00439808 - 0.09346903
Jul8200a <-	0.00194222 -	0.00439808

```

```{r,echo=FALSE , warning=FALSE, message=FALSE}
Julplot <- totalnames %>% filter(str_detect(name, "^Jul")) %>% group_by(year) %>% filter( year < 2018)
Juliaplot <- totalnames %>% filter(str_detect(name, "^Julia$")) %>% group_by(year)  %>% filter(year < 2018) 

ggplot(mapping = aes(x = year, y = prop)) +
  geom_jitter(data = Julplot, color = "blue") + 
  geom_jitter(data = Juliaplot, color = "red") +
  labs(title = "Proportion of names that start with JUL by year", y = "Proportion of Names", subtitle = "Blue is names that start with Jul, Red is Julia")

  Julplot <- totalnames %>% filter(str_detect(name, "^Jul")) %>% group_by(year) %>% filter( year == "2018") %>% summarise(sumprop = sum(prop))
Juliaplot <- totalnames %>% filter(str_detect(name, "^Julia$")) %>% group_by(year)  %>% filter(year == "2018") %>% summarise(sumprop = sum(prop))

ggplot(mapping = aes(x = sumprop)) +
  geom_histogram(data = Julplot, fill = "blue") + 
  geom_histogram(data = Juliaplot, fill = "red") +
  labs(title = "Proportion of names that start with JUL in 2018", x = "Proportion of Names", subtitle = "Blue is names that start with Jul, Red is Julia")
```

```{r,echo=FALSE , warning=FALSE, message=FALSE}
Julplot2018 <- totalnames %>% filter(str_detect(name, "^Jul")) %>% group_by(year) %>% filter(year >= 2000 & year <= 2018) %>% summarise(sumprop = sum(prop))
Julia1plot <- Julia1 %>% group_by(year) %>% filter(year >= 2000 & year <= 2018) %>% summarise(sumprop = sum(prop))

ggplot(mapping = aes(x = year, y = sumprop)) +
  geom_point(data = Julplot2018, color = "blue") +
  geom_point(data = Julia1plot, color = "red") +
  labs(title = "Proportion of names that start with JUL by year from 2000-2018", y = "Proportion of Names", subtitle = "Blue is names that start with Jul, Red is Julia (or an alternate spelling)")

  Julplot8200 <- totalnames %>% filter(str_detect(name, "^Jul")) %>% group_by(year) %>% filter(year >= 1982 & year <= 2000) %>% summarise(sumprop = sum(prop))
Julia1plot <- Julia1 %>% group_by(year) %>% filter(year >= 1982 & year <= 2000) %>% summarise(sumprop = sum(prop))

ggplot(mapping = aes(x = year, y = sumprop)) +
  geom_point(data = Julplot8200, color = "blue") +
  geom_point(data = Julia1plot, color = "red") +
  labs(title = "Proportion of names that start with JUL by year from 1982-2000", y = "Proportion of Names", subtitle = "Blue is names that start with Jul, Red is Julia(or an alternate spelling)")
```

***Regular expression***
I used "^J[eu]+[ew]*l[iey]+a$"  because I thought it would encompass all the different spellings of Julia.

***Name now Compared to when I was Born***

My name is less popular now than it was when I was born. One reason for this is that Romeo and Juliet came out in 1996, four years before I was born. Julia has been a popular name for a while however, was more popular in the late 19th century and early 20th century than it is now. 

***How Popular is your name compared to other names that start with "Jul"?*** 

Julia is relatively popular compared to other names starting with Jul however, Julie was significantly more popular in the 1950s to the 1980s than Julia was. Additionally, today, Julian is more popular than Julia. Overall, Julia is relatively popular because there has only been two names, starting with Jul, that were more popular than Julia.  

## Declan Franklin

## Thomas Neal 

## Thomas Neal 

***Regular expression***
To make all my comparisons I wanted to include both spellings of my name: Thomas and Tomas. To do this I used the regular expression "^[T][h]*omas". Then I compared it to all male names starting with "Tho" and "Tom".

***Plots***

```{r, echo=FALSE , warning=FALSE, message=FALSE}


Thomasall <- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter( sex=="M") %>% group_by(year) %>% summarise(Thomas=sum(n))

Thoall <- totalnames %>% group_by(year) %>%  filter(str_detect(name, "^Tho") | str_detect(name,"^Tom")) %>% filter(sex=="M") %>% summarise(Total=sum(n))


Thomas <- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter(year>1997 & sex=="M") %>% group_by(year) %>% summarise(Thomas=sum(n))



Tho <- totalnames %>% filter(year>1997) %>% group_by(year) %>%  filter(str_detect(name, "^Tho") | str_detect(name,"^Tom")) %>% filter(sex=="M") %>% summarise(Total=sum(n))


comb<- left_join(Thomas,Tho)

comball <- left_join(Thomasall,Thoall)



```

```{r, echo=FALSE , warning=FALSE, message=FALSE}
ggplot(comball) + geom_line(mapping = aes(x=year,y=Total,color="red")) + geom_line(mapping = aes(x=year,y=Thomas,color="blue")) + labs(title= "Popularity of Thomas/Tomas vs all names starting with Tho/Tom (1880-2018)" , x="Year", y="Total babies" , caption = "Black line = Thomas Edison's death" ) +scale_color_discrete(name="Key", labels=c("Thomas/Tomas","Tho/Tom")) + geom_vline(xintercept =1931)

```




```{r,echo=FALSE , warning=FALSE, message=FALSE}

ggplot(comb) + geom_line(mapping = aes(x=year,y=Total,color="red")) + geom_line(mapping = aes(x=year,y=Thomas,color="blue")) + labs(title= "Popularity of Thomas/Tomas vs all names starting with Tho/Tom (1998-2018)" , x="Year", y="Total babies"  ) +scale_color_discrete(name="Key", labels=c("Thomas/Tomas","Tho/Tom")) 


```

```{r,echo=FALSE , warning=FALSE, message=FALSE}
Thomasx <- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter(year>1976 & sex=="M") %>% group_by(year) %>% summarise(Thomas=sum(n))

Thox <- totalnames %>% filter(year>1976) %>% group_by(year) %>%  filter(str_detect(name, "^Tho") | str_detect(name,"^Tom")) %>% filter(sex=="M") %>% summarise(Total=sum(n))

combx <- left_join(Thomasx,Thox)



ggplot(combx) + geom_line(mapping = aes(x=year,y=Total,color="red")) + geom_line(mapping = aes(x=year,y=Thomas,color="blue")) + labs(title= "Popularity of Thomas/Tomas vs all names starting with Tho/Tom(1977-2018)" , x="Year", y="Total babies"  ) +scale_color_discrete(name="Key", labels=c("Thomas/Tomas","Tho/Tom"))
                              

```

***Changes in popularity for my name from 1880 to 2018***

#### At the start:

Thomas begins as a relevantly uncommon name, however, my theory is that this changed when Thomas Jefferson (1847-1931) became a household name. Since he was famous for being brilliant there became a positive association with the name Thomas. This leads the name to skyrocket in popularity from 1930-1950. 

####  When I was born to present:

When I was born Thomas was already on a steep decline and would continue to decline until the present. I think much of this decline is due to the name becoming too popular and thus becoming undesirable. Things such as "Thomas The Train" and actors like Tom Hanks and Tom Crews have disincentivized people who want their kids to have a unique name.

***Popularity comparison***

The name Thomas is the majority of names starting with "Tho" or "Tom". The exception are names like Tom, Thorton, and Tommy. The largest point of diversion was around 1960 when the name "Tom" was exceptionally popular. 




## Nisia Pinto


```{r,echo=FALSE, comment=FALSE, message=FALSE,results=FALSE}
nis<-popnames%>%
  filter(str_detect(name,"^[N]is")) %>% #kicks out a vector of T and F
  filter(year<=2018&year>=1880)%>%
  filter(sex=='F') %>%
  group_by(year)%>%
  summarise(var=n(),sum(count))
nis2<-popnames %>% 
  group_by(year) %>%
  summarise(tot=sum(count))
nis3<-full_join(nis,nis2)%>%
  mutate(prop=`sum(count)`/tot)

nisia<-popnames%>%
  filter(str_detect(name,"^[N]i+[csz][s]*[h]*[iea]*a")) %>%
  filter(year<=2018&year>=1880)%>%
  filter(sex=='F') %>%
  group_by(year)%>%
  summarise(var=n(),sum(count))
nisia2<-popnames %>% 
  group_by(year) %>%
  summarise(tot=sum(count))
nisia3<-full_join(nisia,nisia2)%>%
  mutate(prop=`sum(count)`/tot)
```

```{r,echo=FALSE, comment=FALSE, message=FALSE,warning=FALSE, fig.width=10}
ggplot(nis3)+
    geom_line(aes(x=year,y=prop),color='red')+
  ylab('Proportions')+
  xlab('Year')+
  geom_line(data=nisia3,aes(x=year,y=prop),color='orange')+
  xlim(1950,2018)+
  ggtitle("Proportions of First Three Letters (Red) and Variations of Nisia (Orange) Vs Time") + 
  theme_bw()
```

***Regular expression***
Nisia is a name not found in the dataset.The regular expression I used was ^[N]i+[csz][s]*[h]*[iea]*a$.The first letter I used is N and after I used 'i.' Without an 'i,' it is not a variation of the name Nisia. Additionally, I used the letters c,s,z according to how people pronounce the name Nisia in different situations. I added an extra 's' in cases of Nissa or 'h' in Nisha's case. Lastly, I said that it must include or not 'iea,' but it has to end with an a.

***Name now Compared to when I was Born***
The popularity of my name decreased since the year I was born. Nisia and variations of Nisia are not familiar names.  

***How Popular is your name compared to other names that start with "Nis"?*** 
 According to the plot, my name is not as popular as the names beginning with Nis. The popularity of my name decreased more after the year 2000 compared to its three-letter initial.



# The Little Mermaid Effect 



```{r,echo=FALSE , warning=FALSE, message=FALSE}

arial <- totalnames  %>% filter(str_detect(name, "^A[r]+[iyea][ea]l")) %>% filter(year %in% list(1973, 1988, 1990)) %>% filter(sex=="F") %>% group_by(year) %>% summarise(Unique = n())
arial
```

```{r,echo=FALSE , warning=FALSE, message=FALSE}
rachel <- totalnames %>% filter(str_detect(name, "^R[ae][aey]*[sc]h[aey]+[l]+e?")) %>% filter(year %in% list(1973, 1988, 1990)) %>% filter(sex=="F") %>% group_by(year) #%>% summarise(Unique = n())
rachel
```

```{r,echo=FALSE , warning=FALSE, message=FALSE}
arialAllYears <- totalnames  %>% filter(str_detect(name, "^A[r]+[iyea][ea]l")) %>%filter(sex=="F") %>% group_by(year) %>% summarise(Total=sum(n))
arialAllYears
ggplot(arialAllYears) + geom_line(mapping=aes(x=year,y=Total)) + geom_vline(xintercept =1989, color="red") + labs(title = "Babys born with some form of the name Ariel", x="Year", y="Babies born with a variation of Ariel", caption = "Red line= Release of the Little Mermaid movie")
```



```{r,echo=FALSE , warning=FALSE, message=FALSE}

vowelnames <- totalnames %>% filter(str_detect(name, "^[AEIOU]"),year>=1988 & year <=1990, sex=="F") %>% group_by(year) %>% summarise(otherprop=sum(prop))


ariel19 <-  totalnames  %>% filter(str_detect(name, "^A[r]+[iyea][ea]l"),year>=1988 & year <=1990, sex=="F") %>% group_by(year) %>% summarise(totalprop=sum(prop))

vowelnames
ariel19

compair <- left_join(vowelnames,ariel19)

compair

```

# Your Names 

## Andres Acevedo



***Change in Popularity from 1996 to  2018***

The name Andres was actually growing in popularity throughout much of the 20th century. This could be explained by the increase in immigration from Spain and Central America. The popularity of my name peaked around 1984 and has been declining ever since. The actual change in proportion from 1996 to 2018 was 0.00339 - 0.00923 = -0.00584

***Change in Population from 1974 to 1996***

While Andres has never been a common name, for much of the 20th century, it was on the rise in terms of it's popularity. This continued until around 1984 and has been on the decline ever since. The actual change in proportion from 1974 to 1996 was 0.00923 - 0.00669 = 0.00254. These changes are small and I don't believe they could be categorized as significant.

***Relative Change of The Name***


## Julia Foley 


***Change in Popularity from 2000 to 2018*** 

##### Actual Change in Proportion 

In 2000, the proportion of people named Julia was 0.00439808. In 2018, the proportion of people named Julia was 0.09346903. Therefore, the proportion of people named Julia increased 8.91% from 2000 to 2018.

##### Relative Change in Proportion 

In 2000, the proportion of people named Julia (or some alternate spelling) was .00443067. In 2018, the proportion of people named Julia (or some alternate spelling) was .093625. Therefore, the proportion of people named Julia increased 8.92% from 2000 to 2018. 

***Change in Popularity from 1982 to 2000***

##### Actual Change in Proportion 

In 1982, the proportion of people named Julia was 0.00194222. In 2000, the proportion of people named Julia was 0.00439808. Therefore, the proportion of people named Julia increased .246% from 1982 to 2000. 

##### Relative Change in Proportion 

In 1982, the proportion of people named Julia (or some alternate spelling) was .0019455. In 2000, the proportion of people named Julia (or some alternate spelling) was .00443067. Therefore, the proportion of people named Julia increased .249% from 1982 to 2000. 

**Plots** 

```{r}
Julplot2018a <- Julplot2018 %>% mutate(Year = year == "2000" | year == "2018")

ggplot(data = Julplot2018a, mapping = aes(x = sumprop, fill = Year)) +
  geom_histogram() +
  labs(title = "Proportion of people named Julia (or an alternate spelling) by year from 2000-2018", y = "Number of Years with Proportion", x = "The sum of the proportion of people named Julia(or an alternate spelling)", subtitle = "Blue is 2000 and 2018, Red is all other names")
```

```{r}
Julplot8200a <- Julplot8200 %>% mutate(Year = year == "1982" | year == "2000")

ggplot(data = Julplot8200a, mapping = aes(x = sumprop, fill = Year)) +
  geom_histogram() +
  labs(title = "Proportion of people named Julia (or an alternate spelling) by year from 1982-2000", y = "Number of Years with Proportion", x = "The sum of the proportion of people named Julia(or an alternate spelling)", subtitle = "Blue is 1982 and 2000, Red is all other names")
```

## Declan Franklin

## Thomas Neal 


```{r, echo=FALSE, message=FALSE, warning=FALSE}
Thomas1997 <- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter( sex=="M" & year==1997)  %>%  summarise(Thomas=sum(prop))
Thomaswork

Thomas2018 <- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter( sex=="M" & year==2018)  %>%  summarise(Thomas=sum(prop))



```
***Change in Popularity from 1998 to 2018*** 
In 1998 the proportion of people named Thomas or Tomas was:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
Thomas1997 <- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter( sex=="M" & year==1997)  %>%  summarise(Thomas=sum(prop))
Thomaswork
```

In 2018 the In 1998 the proportion of people named Thomas or Tomas was:

```{r, echo=FALSE, message=FALSE, warning=FALSE}


Thomas2018 <- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter( sex=="M" & year==2018)  %>%  summarise(Thomas=sum(prop))
Thomas2018

```

So the change in proportion from 1998 to 2018 is:

```{r, echo=FALSE, message=FALSE, warning=FALSE}


Thomasprop<-Thomas2018$Thomas-Thomas1997$Thomas
Thomasprop

```

***Change in Popularity from 1998 to 2018*** 
In 1977 the proportion of people named Thomas or Tomas was:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
Thomas1977<- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter( sex=="M" & year==1977)  %>%  summarise(Thomas=sum(prop))
Thomaswork
```

In 1998 the proportion of people named Thomas or Tomas was:

```{r, echo=FALSE, message=FALSE, warning=FALSE}


Thomas1998 <- totalnames %>% filter(str_detect(name,"^[T][h]*omas$")) %>% filter( sex=="M" & year==1998)  %>%  summarise(Thomas=sum(prop))
Thomas1998

```

So the change in proportion from 1977 to 1998 is:

```{r, echo=FALSE, message=FALSE, warning=FALSE}


Thomasprop<-Thomas1998$Thomas-Thomas1977$Thomas
Thomasprop

```

## Nisia Pinto

```{r,echo=FALSE, comment=FALSE, message=FALSE,warning=FALSE}
L1<- nisia3 %>% filter(year %in% 1995) %>% group_by(year) %>% summarise(prop_A = sum(prop))  
L2<- nisia3 %>% filter(year %in% 2018) %>% group_by(year) %>% summarise(prop_A = sum(prop)) 
L3<-nisia3 %>% filter(year %in% 1972) %>% group_by(year) %>% summarise(prop_A = sum(prop)) 
Change1<-1.777853e-05-5.653651e-05		 #L2-L1
Change2<-5.653651e-05-3.27647e-05
Rchange<-Change1/5.653651e-05
Rchange2<-Change2/3.27647e-05
```


***Change in Popularity from 1995 to 2018*** 
In 2018, the proportion of people named variation of Nisia was 1.777853e-05 and in 1995 was    5.653651e-05. Therefore, there was a negative change of 3.875798e-05, showing that the name became less popular.

***Change in Popularity from 1972 to 1995***
In 1995, the proportion of people named Leonela was 5.653651e-05 and in 1972 was 3.27647e-05. Therefore, there was a  change of  2.377181e-05, showing that the name increased in popularity.


***Relative Change of The Name***
Nisia had a relative change of -0.685539 between 1995 to 2018 and a corresponding relative change of 0.7255311 between 1972 to 1995.


```{r,echo=FALSE, comment=FALSE, message=FALSE,warning=FALSE}
nisia4<-nisia3 %>% filter(year>=1995)

ggplot(nisia4,aes(prop))+
  geom_histogram(color='orange')+
  ggtitle("Proportions of the name Nisia 1995 to 2018") + 
  xlab("Proportions")+ 
  ylab('Dataset')+
  theme_bw()
```
```{r,comment=FALSE, message=FALSE,warning=FALSE}
nisia5<-nisia3 %>% filter(year<=1995,year>=1972)

ggplot(nisia5,aes(prop))+
  geom_histogram(color='red')+
  ggtitle("Proportions of the name Nisia 1995 to 1972") + 
  xlab("Proportions")+ 
  ylab('Dataset')+
  theme_bw()
```


## Team 

# Name Recommendation 
```{r,echo=FALSE , warning=FALSE, message=FALSE}
yrname <- babynames %>% group_by(name) %>% count(name)
yr2017 <- babynames %>% filter(year == "2017")
yr2016 <- babynames %>% filter(year == "2016")
yr2015 <- babynames %>% filter(year == "2015")
yr2014 <- babynames %>% filter(year == "2014")
yr2013 <- babynames %>% filter(year == "2013")


Lee1 <- yr2017 %>% filter(name == "Lee")
yr20181 <- yob2018 %>% arrange(n) %>% head(31933)
common2018 <- yr20181 %>% count(name)
common22018 <- common2018 %>% filter(n == "2")
common32018 <- common22018 %>% left_join(yr20181, by = "name") 
common42018 <- common32018 %>% group_by(name) %>% summarise(nsum = sum(n.y))
common52018 <- common32018 %>% filter(gender == "M") %>% select(name, n.y)
common62018 <- common42018 %>% left_join(common52018, by = "name") %>% mutate(per = n.y/nsum) 
common72018 <- common62018 %>% filter(per < .75 & per > .25)
common82018 <- yr20181 %>% anti_join(common72018, by = "name")

```

```{r,echo=FALSE , warning=FALSE, message=FALSE}
trend1 <- common82018 %>% left_join(babynames, by = "name") %>% filter(!is.na(n))
trend2 <- trend1 %>% group_by(name) %>% summarise(meantot = mean(prop))
trend1716 <- trend1 %>% filter(year == "2016" | year == "2017") %>% group_by(name) %>% summarise(mean17 = mean(prop))
trend1516 <- trend1 %>% filter(year == "2016" | year == "2015") %>% group_by(name) %>% summarise(mean16 = mean(prop))
trend3 <- trend1716 %>% left_join(trend2, by = "name") %>% mutate(trend = mean17/meantot)
trend4 <- trend1716 %>% left_join(trend1516, by = "name") %>% mutate(trend = mean16/mean17) %>% filter(trend > 1)
trend1514 <- trend1 %>% filter(year == "2014" | year == "2015") %>% group_by(name) %>% summarise(mean15 = mean(prop))
trend5 <- trend4 %>% left_join(trend1514, by = "name") %>% mutate(trend = mean15/mean16) %>% filter(!is.na(trend)) %>% filter(trend > 1)
trend1314 <- trend1 %>% filter(year == "2014" | year == "2013") %>% group_by(name) %>% summarise(mean14 = mean(prop))
trend6 <- trend5 %>% left_join(trend1314, by = "name") %>% mutate(trend = mean14/mean15) %>% filter(!is.na(trend)) %>% filter(trend > 1)
trend1312 <- trend1 %>% filter(year == "2012" | year == "2013") %>% group_by(name) %>% summarise(mean13 = mean(prop))
trend7 <- trend6 %>% left_join(trend1312, by = "name") %>% mutate(trend = mean13/mean14) %>% filter(!is.na(trend)) %>% filter(trend > 1) 
trend8 <- trend7 %>% select(name)
```

```{r,echo=FALSE , warning=FALSE, message=FALSE}
nick1 <- trend8 %>% left_join(yob2018, by = "name") 
nick2 <- nick1 %>% filter(gender == "M")
nick3 <- nick1 %>% filter(gender == "F")
nick4 <- nick2 %>% summarise(mean(num))
nick5 <- nick3 %>% summarise(mean(num))
nick8 <- nick2 %>% filter(num > 419)
nick9 <- nick3 %>% filter(num > 282)
nick8 <- nick8 %>% arrange(desc(num)) %>% head(50)
nick9 <- nick9 %>% arrange(desc(num)) %>% head(50)
finalgroup <- nick8 %>% left_join(trend3) %>% arrange(trend)
finalgroup1 <- nick9 %>% left_join(trend3) %>% arrange(trend)
```
## Recommendation 

We recommend Kevin for a boy and Michelle for a girl.

## Why We Recommended those Names

We recommend this because these are both that are clearly for one gender. Additionally, they are names that do not have automatic nicknames associated with them. Additionally, Kevin and Michelle are both trending however, are not overly trendy. Both of these names are good names because they represent the criteria that the client gave us and their unique enough that not every kid will have these names. 

## Plots for Recommendation 

```{r}
avgpropmen <- totalnames %>% group_by(year) %>% filter(sex == "M") %>% mutate(average = cummean(prop))
avgpropwomen <- totalnames %>% group_by(year) %>% filter(sex == "F") %>% mutate(average = cummean(prop))
propkeith <- totalnames %>% filter(name == "Keith")
propmichelle <- totalnames %>% filter(name == "Michelle") 
ggplot(mapping = aes(x = year, y = prop)) + 
  geom_point(data = avgpropmen, color = "blue") +
  geom_point(data = propkeith, color = "red")
```

```{r}
ggplot(mapping = aes(x = year, y = prop)) + 
  geom_point(data = avgpropwomen, color = "blue") +
  geom_point(data = propmichelle, color = "red")
```







### Thomas Neal

I tidied and graphed the data for the Little Mermaid. I also helped to figure out how to set up the data frame to best graph our individual names over time. I calculated the proportions for the additional 2018 data. I also did all of my individual parts and helped to format the lab.
