---
title: "Lab12_P-Hackers"
author: "Declan Franklin"
date: "11/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(gapminder)
library(tidyverse)
```


# Two-Sample Permutation Test

Say we want to know if groups $A$ and $B$ have different means for some variable like age or gdp per capita. We conduct a hypothesis test where the null hypothesis $H_0$ is that the means are the same and the alternative hypothesis $H_1$ that the means are different. 

> **H0:** $\mu_A - \mu_B = 0$  
**H1:** $\mu_A - \mu_B \neq 0$

To test our hypothesis, we run an experiment in which we take samples (of age, gdp per capita, etc.) of sizes $n_A$ and $n_B$ from populations $A$ and $B$. Then we calculate the sample means of both groups $\bar{X_A}$ and $\bar{X_B}$.

If the null hypothesis is true, $\bar{X_A} - \bar{X_B} \approx 0$.

We mix up the group labels $A$ and $B$ over our sampled data many times. We compute $\bar{X_A} - \bar{X_B}$ fr each iteration and make a histogram of all the iterations. We find the actual $\bar{X_A} - \bar{X_B}$ percentile on the histogram. We reject the null hypothesis in favor of the alternative if the percentile of the actual $\bar{X_A} - \bar{X_B}$ is in the critical region.


You will implement the following function, perm_mean, to create 1000 iterations of the two-sample permuation test. Fill in the function-skeleton to implement the function.

```{r}
mena_ans <- gapminder %>%
  filter(continent== "Asia" | continent == "Africa") %>%
  select(continent,lifeExp) 
```


```{r, eval = F}

gpa <- c(3.0,2.6,2.0,3.0,3.4)



perm_mean <- function(perms = 1000, all_values, n_A)  # All values of hist and cmci
{
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  perm_mean_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "all values" into disjoint 
    # groups of size "nA" and "nB" respectively
    # we will take the vector, shulf it, and call n_A of the top and call it group A, then the rest will be group B
    rand_order <- sample(all_values)
    group_A <- rand_order[1:n_A] # First go to n_A
    group_B <- rand_order[n_A:length(all_values)] # Rest go here
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    
    A_bar <- mean(group_A)
    B_bar <- mean(group_B)
    print(group_A)
   
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1

    diff_means <- A_bar - B_bar
    perm_mean_diffs[i] <- diff_means
  }
  
  # Step 5:
  # Return the permuted mean differences vector
return(perm_mean_diffs)
  
}
```

```{r}
perm_mean(100,mena_ans$lifeExp,5)
```



# Linear Correlation Permuation Test

Say we want to figure out if two variables in a dataset are linearly correlated. For example, is a person's age correlated with a person's income?

Recall that the Pearson correlation coefficient measures the strength of the linear relationship $\rho$ between two random variables, $X$ and $Y$. A correlation close to 1 indicates a strong, positive linear relationship. A correlation close to -1 indicates a strong, negative linear relationship. A correlation close to 0 indicates no linear relationship.

Our null hypothesis is that the variables are not correlated. Our alternative hypothesis is that the variables are correlated.

> **H0:** $\rho = 0$  
**H1:** $\rho \neq 0$

To test our hypothesis, we shuffle the labels on both variables many times. For each iteration, we compute the Pearson linear correlation $\rho$. Make a histogram of all the shuffled correlations. Plot the actual correlation between the variables on the histogram. Reject the null hypothesis if the percentile of the actual correlation is in the critical region.


Implement the following function, perm_cor, to create 1000 iterations of the correlation permuation test. Fill in the function-skeleton to implement the function.

```{r,echo=FALSE,message=FALSE}
americA <- gapminder %>%
  filter(country== "United States", year > 1952, year < 1977)

cubA <- gapminder %>%
  filter(country== "Cuba", year > 1952, year < 1977)

americAfter <- gapminder %>%
  filter(country== "United States", year > 1977, year < 2002)

cubAfter <- gapminder %>%
  filter(country== "Cuba", year > 1977, year < 2002)

  
```


```{r, eval = F}

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
   perm_mean_diffs2 <- numeric(perms)
  
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    rand_order_y <- sample(y)
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    cor_comp <- cor(x,rand_order_y)
    perm_mean_diffs2[i] <- cor_comp 
    
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(perm_mean_diffs2)
}
```




### 2. Is there a strong corrilation of gdpPercap in Cuba or America during the space race? What about decades after?

<p>

#### Reject the null if the percentile is less then 95%


**The Histogram during the Space Race**
```{r,echo=FALSE}
twohist <- hist(perm_cor(1000,americA$gdpPercap,cubA$gdpPercap),main = "GDP/Capita in America and Cuba during the Cold War", xlab = "Linear Correlation")

twohist 

```

**The Percentiles during the Space Race**
```{r, echo=FALSE, message=FALSE, warning=FASLE}

twopen <- quantile(perm_cor(1000,americA$gdpPercap,cubA$gdpPercap),0.95) 

twopen
```


### Decades after the space race
<p>

#### Reject th null if the perceentile is less then 95%

**The Histogram After the Space Race**
```{r,echo=FALSE}
afterhist <- hist(perm_cor(1000,americAfter$gdpPercap,cubAfter$gdpPercap),main = "GDP/Capita in America and Cuba After the Cold War", xlab = "Linear Correlation")

afterhist 

```
**The Percentiles After the Space Race**
```{r, echo=FALSE, message=FALSE, warning=FASLE}

afterpen <- quantile(perm_cor(1000,americAfter$gdpPercap,cubAfter$gdpPercap),0.95) 

afterpen
```



#### Conclussion 
<p>
In conclusion, during the space race, there was a corrilation betwen GDP/capita. the percentile was at 95.1%, which agrees with our null hypothesis. After the space race, the percentile was well below our null hypothesis, so we can't conclude a corrilation. 



# Gapminder Questions

Your data science team must come up with two specific questions of interest about the gapminder data set (from 'gapminder' library). 

1.Is life expectancy longer in Africa than in Asia? 

2. Is there a strong corrilation of gdpPercap in Europe or America during the space race? What about a decade after?


Gdp> Pre and and war 





# Briefly describe the gapminder dataset.

The gapminder data set contains data between 1952 and 2007. The data includes Countries, Continents, Year, Life Expectancy, Population, and GDP per Capita. Within this data, we can analyze data about the relationship within specific countries and the relationship between countries. 


# Why these questions are important

1. With Asia having the largest population in the world, and Africa being one of the richest in natural resources it is interesting to compare their life expextancy. Asia has historically dealt with overpopulation and the problems this presents when trying to keep people healthy. Africa has been exploited for its resources and thus has had to trouble maintaining healthy standards for its people. By looking at comparisons in their data we may be able to find a correlation between their struggles.

2. With the space race representing the race between the two economic ideologies of the time, capitalism and communism it is important to see how these two ideas impacted surrounding countries' economies. By comparing the GDPs we can gain a sense of the impact these to contradicting ideas had.






# Format and Order for Team Lab Report

Create a team R Markdown document knitted to html and submitted to OSF. Put your team name in the filename. Make sure your code works on the updated dataset realeased on Friday.

# What each individual do 

### Julia Foley

We got together Wednesday night to work on the lab. We did the labe together however, I helped with the Linear Correlation Permulation Test, the questions of the lab, and the description of the dataset.  

### Andres Acevedo 

<p>
### Declan Franklin
<p>

I was able to work alongside Thomas, Nisia, and Julia to figure out the linear correlation function. After completed said function, I used to say function to answer our second question about the space race.


### Thomas Neal 

### Nisia Pinto

We met on Wedsnesday to go over the questions in the lab. I helped braistorming the questions and helped in the linear correlation function.



