---
title: "P-HackersLab12"
author: "Team P-Hackers"
date: "11/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
library(gapminder)
library(tidyverse)
```

# Description of gapminder Dataset.

The gapminder data set contains data between 1952 and 2007. The data includes Countries, Continents, Year, Life Expectancy, Population, and GDP per Capita. Within this data, we can analyze data about the relationship within specific countries and the relationship between countries.


# Gapminder Questions

Your data science team must come up with two specific questions of interest about the gapminder data set (from 'gapminder' library). 

1. Is life expectancy longer in Africa than in Asia? 

2. Is there a strong corrilation of gdpPercap in Europe or America during the space race? What about a decade after?


# Why these questions are important

1. With Asia having the largest population in the world, and Africa being one of the richest in natural resources it is interesting to compare their life expextancy. Asia has historically dealt with overpopulation and the problems this presents when trying to keep people healthy. Africa has been exploited for its resources and thus has had to trouble maintaining healthy standards for its people. By looking at comparisons in their data we may be able to find a correlation between their struggles.

2. With the space race representing the race between the two economic ideologies of the time, capitalism and communism it is important to see how these two ideas impacted surrounding countries' economies. By comparing the GDPs we can gain a sense of the impact these to contradicting ideas had.



```{r,,echo=FALSE,message=FALSE}
mena_ans <- gapminder_unfiltered %>%
  filter(continent== "Asia" | continent == "Africa") %>%
  select(continent,lifeExp) 
```


```{r, eval = T,echo=FALSE,message=FALSE}
perm_mean <- function(perms = 1000, all_values, n_A)  # All values of hist and cmci
{
  ## Variables ##
  # perms: The number of permutations 
  # all_values (num): all data values
  # n_A (int): Size of group A
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
  perm_mean_diffs <- numeric(perms)
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly separate vector "all values" into disjoint 
    # groups of size "nA" and "nB" respectively
    # we will take the vector, shulf it, and call n_A of the top and call it group A, then the rest will be group B
    rand_order <- sample(all_values)
    group_A <- rand_order[1:n_A] # First go to n_A
    group_B <- rand_order[n_A:length(all_values)] # Rest go here
    # Step 3:
    # Compute the sample means for the two groups from 
    # step 2
    
    A_bar <- mean(group_A)
    B_bar <- mean(group_B)
    
   
    
    # Step 4: 
    # Compute the difference in sample means, store the
    # value in the vector from step 1

    diff_means <- A_bar - B_bar
    perm_mean_diffs[i] <- diff_means
  }
  
  # Step 5:
  # Return the permuted mean differences vector
return(perm_mean_diffs)
  
}
```



```{r,echo=FALSE,message=FALSE}
americA <- gapminder_unfiltered %>%
  filter(country== "United States", year >= 1955, year <= 1957)

russiA <- gapminder_unfiltered %>%
  filter(country== "Russia", year >= 1950, year <= 1990)

americAfter <- gapminder_unfiltered %>%
  filter(country== "United States", year >= 1991, year <= 2002)

russiAfter <- gapminder_unfiltered %>%
  filter(country== "Russia", year >= 1991, year <= 2002)

  
```


```{r, eval = T,echo=FALSE,message=FALSE}

perm_cor <- function(perms = 1000, x, y)
{
  ## Variables ##
  # perms: The number of permutations 
  # x: Vector of Variable 1 - for computing correlation
  # y: Vector of Variable 2 - for computing correlation
  ###############
  
  # Step 1:
  # Create vector of zeroes of length "perms" to store
  # permuted mean differnces
   perm_mean_diffs2 <- numeric(perms)
  
  
  # Loop throught number of permutations
  for (i in c(1:perms))
  {
    # Step 2:
    # Randomly mix up the values in the vector "y"
    rand_order_y <- sample(y)
    # Step 3:
    # Compute the correlation between x and the randomly mixed
    # up y-vector. Store this value in the vector from step 1.
    cor_comp <- cor(x,rand_order_y)
    perm_mean_diffs2[i] <- cor_comp 
    
  }
  
  # Step 4:
  # Return new updated vector, created in step 1
  return(perm_mean_diffs2)
}
```


### 1. Is life expencity longer in Africa than in Asia?

#### Accept the the null if the p-value less than .05

**Histogram of life expediency between Asia and Africa**

```{r,echo=FALSE,message=FALSE}
hist2 <-hist(perm_mean(100,mena_ans$lifeExp,5),main = "Distrodution of mean diffrence in life expectancy", xlab = "Linear Correlation")
hist2
```
**The 95 Percentile for Africa vs Asia**

```{r,echo=FALSE}
question1pen <- quantile(perm_mean(100,mena_ans$lifeExp,5),0.95)
question1pen
```

#### Conclusion To Question 1
We found that the 95th Percentile of mean differences was 0.9328. This suggests that we should accept our null hypothesis and state that we have not proven a statistically significant difference in the life expectancy of Asain countries and African countries.


### 2. Is there a strong corrilation of gdpPercap in Russia and America during the space race? What about decades after?

<p>

#### Reject the null if the percentile is less then 95%


**The Histogram during the Space Race**
```{r,echo=FALSE}
twohist <- hist(perm_cor(1000,americA$gdpPercap,russiA$gdpPercap),main = "GDP/Capita in America and Russia during the Cold War", xlab = "Linear Correlation")

twohist 

```

**The Percentiles during the Space Race**
```{r, echo=FALSE, message=FALSE, warning=FALSE}

twopen <- quantile(perm_cor(1000,americA$gdpPercap,russiA$gdpPercap),0.95) 

twopen
```

### Decades after the space race
<p>

#### Reject the null if the percentile is less then 95%

**The Histogram After the Space Race**
```{r,echo=FALSE}
afterhist <- hist(perm_cor(1000,americAfter$gdpPercap,russiAfter$gdpPercap),main = "GDP/Capita in America and Russia After the Cold War", xlab = "Linear Correlation")

afterhist 

```
**The Percentiles After the Space Race**
```{r, echo=FALSE, message=FALSE, warning=FALSE}

afterpen <- quantile(perm_cor(1000,americAfter$gdpPercap,russiAfter$gdpPercap),0.95) 

afterpen 
```



#### Conclusion to Question 2
<p>
In conclusion, during the space race, there was a corrilation betwen GDP per-capita. the percentile was 48.94%, which suggests that we should accept our null hypothesis. After the space race, the percentile was well below our null hypothesis, so we can't conclude the variables are corrilated. 



# Individual Contributions 

### Julia Foley

We got together Wednesday night to work on the lab. We did the labe together however, I helped with the Linear Correlation Permulation Test, the questions of the lab, and the description of the dataset.  

### Andres Acevedo

Unfortunately, I wasn't available to meet with my team to formulate the original questions for this lab. However, while going through the team code, I noticed that one of our dataframes contained no information. This had implications for the veracity of our results. I wrote the first function perm_mean which allows us to calculate the first permutation test. I knitted together final document.

### Declan Franklin

I was able to work alongside Thomas, Nisia, and Julia to figure out the linear correlation function. After completed said function, I used to say function to answer our second question about the space race.

### Thomas Neal 

I helped to graph the data for question 1. I also calculated the percentiles for it. I then helped draw the data-based conclusion to this question. Lastly, I helped format the lab. 

### Nisia Pinto

We met on Wedsnesday to go over the questions in the lab. I helped braistorming the questions and helped in the linear correlation function.
