---
title: "Lab 6"
author: "Julia Foley"
date: "10/1/2019"
output: html_document
---

```{r setup, include=FALSE, ECHO = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
COflights <- read_csv("https://files.osf.io/v1/resources/3z5as4pej8/providers/osfstorage/5a8ca28f57103100104584db")
```

```{r, echo=FALSE}
DENFlights <- COflights %>%
  filter(ORIGIN == "DEN" | DEST == "DEN")
```

# Team Research 

<p>

## Question 1 
<p>
**What is the probability that a flight arrives on time in Denver given that it is the holiday season (October - December)?** 

### Hypothesis

**The Holiday season will result in a high percentage of flights arriving late.** 


### First we will declare the events in our sample space. Event E: "the flights that arrive on time in Denver". Event H: "the flights that arrive to Denver during the holiday season". Our conditional probability computation is \[ P(E|S)=\frac{{P(ES)}}{P(E)}\] 


### Next we will filter out flights that arrived on time to denver during the holiday season. This is the numerator of our conditional probability \[ P(E|S)=\frac{\mathbf{P(ES)}}{P(E)}\]
```{r}
numt1 <- COflights %>%
  filter(ARR_TIME<= 14,DEST == "DEN", between(MONTH,10,12)) %>%
  summarise(sum=n())
```
<p>

### Similarly, we will create the denominator that filter all the planes that arrive to Denver during the holiday season. \[ P(E|S)=\frac{P(ES)}{\mathbf{P(E)}}\]

```{r}
demt1 <- COflights %>%
  filter(!is.na(ARR_TIME),DEST == "DEN",between(MONTH,10,12)) %>%
  summarise(sum=n())
```
<p>


### Finally, we can compute our conditional probability \[ P(D|S)=\frac{P(DS)}{P(D)}\] 
```{r}
anst1 <- numt1 / demt1
anst1
```


<p>
### Visual Representation 
```{r}
HolidaySeason <- COflights %>% filter(ARR_DELAY > 15 ,CANCELLED == 0, DEST == 'DEN') 

ggplot(HolidaySeason, mapping = aes(x = MONTH)) + 
geom_histogram(fill = 'steelblue',binwidth = 1, color = 'turquoise2') +
 scale_x_continuous(breaks = seq(0, 12, 1)) + 
 ggtitle('Count of Delayed Flights to Denver by Month') + 
 xlab('Month') + ylab('Number of Flights') + 
 theme_minimal()

```

### Conclusion and Recommendation 

**The data shows that the probability a flight to Denver International Airport will arrive into DIA on time given that it's the holiday season is about 49.1%. This means, during the 2017 holiday season, people travelling to DIA had close to a 50% chance of arriving late (or on time). This could be a problem for connecting flights, or for people trying to catch scheduled rides from the airport. During the holiday season, P-Hackers advice that people account for delays when scheduling rides and for their holiday plans. For DIA, it's advised that they either higher more employees so that they can help with the flow of people, or they could offer deals for cheaper flights at time when less people fly. The deals could incentivize people to take the cheaper flights, and in return, the airport could be less hectic.** 




## Question 2
<p>
**What is the probability that a flight is delayed departing or arriving in Denver given the time of day (specifically 6:00, 12:00, 18:00, and 24:00)?** 

### Why Important for Question 2

Question 2 is important because it could determine what time of day costumers want to travel. It also represents what time of day there are more delays. It is important to study what time of day there is more delays to determine why delays are happening and when the delays are the highest. 

### Deeper Explanation for Question 2

This is important because if they can determine when delays are the highest, they can find why delays are happening. If delays are higher at night, it could be because delays are stacking on top of each other in a domino like effect. However, if delays are higher in the morning, it could be because of something that happened over night. These are not the only explanations to what could have happened, they just show that the time of day impacts why the delays happened. Additionally, the airport could communicate to the citizens what kind of delays they can expect depending on the time of day. 

```{r, echo=FALSE}
team1 <- DENFlights %>%
  filter(ARR_DELAY >= 15 | ARR_DELAY <= -15)
```

### Hypothesis
<p>

**We wanted to compute the probability of a flight being delayed at four times so we chose 6, 12, 18, and 24 respectively. We believe that there will be more flights delayed at 6 and 12 because we believe that there are more flights in the morning than in the evening.**  

```{r, echo=FALSE}
delay2 <- DENFlights %>%
  filter(DEP_TIME == "0600") %>%
  mutate(delay3 = ARR_DELAY >= 15 | ARR_DELAY <= -15)
```
**Probabilities**

The probability that a flight will arrive or depart in Denver is 93.62% 
```{r, echo=FALSE}
count(DENFlights) / count(COflights) * 100
```

###### 6:00

```{r, echo=FALSE}
ggplot(delay2, mapping = aes(x = DEP_TIME, fill = delay3)) + 
  geom_bar() +
  labs(x = "DepartureTime", y = "Number of Flights")
```


The Probability of a flight going in and out of Denver will be delayed at 6:00 is 35.45% 

```{r, echo=FALSE}
team2 <- DENFlights %>%
  filter(DEP_TIME == "0600")
```

```{r, echo=FALSE}
team3 <- team1 %>%
  filter(DEP_TIME == "0600")
```

```{r, echo=FALSE}
count(team3) / count(team2) *100
```

```{r, echo=FALSE}
delay4 <- DENFlights %>%
  filter(DEP_TIME == "1200") %>%
  mutate(delay5 = ARR_DELAY >= 15 | ARR_DELAY <= -15)
```

###### 12:00

```{r, echo=FALSE}
ggplot(delay4, mapping = aes(x = DEP_TIME, fill = delay5)) + 
  geom_bar() +
  labs(x = "DepartureTime", y = "Number of Flights")
```

The Probability of a flight going in and out of Denver will be delayed at 12:00 is 43.51% 

```{r, echo=FALSE}
team7 <- DENFlights %>%
  filter(DEP_TIME == "1200")
```

```{r, echo=FALSE}
team4 <- team1 %>%
  filter(DEP_TIME == "1200")
```

```{r, echo=FALSE}
count(team4) / count(team7) *100
```

```{r, echo=FALSE}
delay6 <- DENFlights %>%
  filter(DEP_TIME == "1800") %>%
  mutate(delay7 = ARR_DELAY >= 15 | ARR_DELAY <= -15)
```

###### 18:00

```{r, echo=FALSE}
ggplot(delay6, mapping = aes(x = DEP_TIME, fill = delay7)) + 
  geom_bar() +
  labs(x = "DepartureTime", y = "Number of Flights")
```

The Probability of a flight going in and out of Denver will be delayed at 18:00 is 47.29% 

```{r, echo=FALSE}
team8 <- DENFlights %>%
  filter(DEP_TIME == "1800")
```

```{r, echo=FALSE}
team5 <- team1 %>%
  filter(DEP_TIME == "1800")
```

```{r, echo=FALSE}
delay8 <- DENFlights %>%
  filter(DEP_TIME == "2400") %>%
  mutate(delay9 = ARR_DELAY >= 15 | ARR_DELAY <= -15)
```

```{r, echo=FALSE}
count(team5) / count(team8) *100
```

###### 24:00

```{r, echo=FALSE}
ggplot(delay8, mapping = aes(x = DEP_TIME, fill = delay9)) + 
  geom_bar() +
  labs(x = "DepartureTime", y = "Number of Flights")
```

The Probability of a flight going in and out of Denver will be delayed at 24:00 is 41.18% 

```{r, echo=FALSE}
team9 <- DENFlights %>%
  filter(DEP_TIME == "2400")
```

```{r, echo=FALSE}
team6 <- team1 %>%
  filter(DEP_TIME == "2400")
```


```{r, echo=FALSE}
count(team6) / count(team9) *100
```


### Conclusion 



**Question 2**
The probabilities show that there is a higher chance of delay at 18:00 then any of the four times we studied. Additionally, there is a lower chance of delay at 6:00 than any of the four times. However, all of the probabilities are relatively similar. 

<p>
### Recommendation 
We recommend for costumers to fly earlier in the day because there is a lower probability of delays at 6:00 then at 18:00.  


<p>
# Individual Research 
<p>

## Julia Foley
<p>

**Question** 

What is the probability a flight has a Late Aircraft Delay and a departure delay over 15 minutes late given that it has a arrival delay over 15 minutes late? 


**Hypothesis** 
I think that there will be a good amount of flights with both departure and arrival delays. However, in relation to the total amount of flights, there will probably be a very small number of flights with all three (departure delay, arrival delay, late aircraft delay). 


**Probabilities and Plots**

```{r, echo=FALSE}
DENFlights1 <- DENFlights %>%
  filter(ARR_DELAY >= 15) 
```

The probability that a flight in Denver will have an arrival delay over 15 minutes late is 17.1%
```{r, echo=FALSE}
count(DENFlights1) / count(DENFlights) *100
```

```{r, echo=FALSE}
DENFlights12 <- DENFlights1 %>%
  filter(DEP_DELAY >= 15) %>%
  filter(LATE_AIRCRAFT_DELAY > 0)
```

The probability that a flight in Denver will have a Late Aircraft delay and a depature delay of over 15 minutes late is 8.43%
```{r, echo=FALSE}
count(DENFlights12) / count(DENFlights) * 100
```

The probability that a flight in Denver will have a Late Aircraft delay and a depature delay of over 15 minutes late given that it has an arrival delay has over 15 minutes late is 49.32%

```{r, echo=FALSE}
count(DENFlights12) / count(DENFlights1) *100
```

The probability that a flight in Denver will have a Late Aircraft Delay, a depature delay over 15 minutes late, and an arrival delay over 15 minutes late is 7.9% 

```{r, echo=FALSE}
count(DENFlights12) / count(COflights) * 100
```

```{r, echo=FALSE}
weather2 <- DENFlights %>%
  filter(!is.na(LATE_AIRCRAFT_DELAY)) %>%
  mutate(weather3 = LATE_AIRCRAFT_DELAY > 0)
```



```{r, echo=FALSE}
ggplot(weather2, mapping = aes(x = DEP_DELAY, y = ARR_DELAY)) +
  geom_point() +
  facet_grid(~weather3) +
  labs(title = "Whether or Not There is a Late Aircraft Delay when Comparing Arrival Delay and Departure Delay", subtitle = "False represents that there is no Late Aircraft Delay, True represents that there is an Late Aircraft Delay", x = "Departure Delay", y = "Arrival Delay") 
```


From the plot, it appears that as the depature delays increase, the arrival delays increase. Additionally, there appears to be a equal amount of Late Aircraft delays to flights that are not Late Aircraft Delays. 
<p>

**Findings** 

I found that out of the flights arriving or departing in Denver, 17.1% of the flights have an arrival time more than 15 minutes late. Of the flights with an arrival time more than 15 minutes late, 49.32% of the flights have a depature time of over 15 minutes late and a late aircraft delay. The fact that the number of flights with a depature time of over 15 minutes late and a late aircraft delay is about half of the flights with an arrival time over 15 minutes late makes sense because the graph shows that about half of the flights have a late aircraft delay and about half do not. Additionally, 8.43% of flights arriving and departing in Denver have an arrival time over 15 minutes late, a departure time over 15 minutes late, and a late aircraft delay. Additionally, 7.9% of flights total have an arrival time over 15 minutes late, a departure time over 15 minutes late, and a late aircraft delay in Denver. This shows that though it is common to have a delayed departure time over 15 minutes late and a late aircraft delay when you have an arrival time over 15 minutes late, it is not very common to have all three out of the total number of flights.  

## Thomas Neal


**Question**

For any given a carrier what is the probability that a flight will be delayed.

**Probabilities**

```{r, echo=FALSE}
car <- COflights %>% filter(CANCELLED!=1) %>% filter(DIVERTED!=1) %>% mutate(Delayed=ARR_DELAY>=15) %>% group_by(CARRIER) 

tot<- car %>% summarise(sum=n()) 
deltot<- car %>% filter(ARR_DELAY>=15) %>% summarise(delsum=n())

Probability <- deltot$delsum/tot$sum

Carriers <- c("AA", "AS", "B6", "DL", "EV", "F9", "NK", "OO", "UA", "VX", "WN")

Probtab <- data.frame(Carriers,Probability)

print(Probtab)
```

**Graph**

Comparing total flights by carrier to delayed flights by carrier.

```{r, echo=FALSE}
car <- COflights %>% filter(CANCELLED!=1) %>% filter(DIVERTED!=1) %>% mutate(Delayed=ARR_DELAY>=15) %>% group_by(CARRIER) 

tot<- car %>% summarise(sum=n()) 
deltot<- car %>% filter(ARR_DELAY>=15) %>% summarise(delsum=n())

Probability <- deltot$delsum/tot$sum

Carriers <- c("AA", "AS", "B6", "DL", "EV", "F9", "NK", "OO", "UA", "VX", "WN")

Probtab <- data.frame(Carriers,Probability)

ggplot(car) +geom_bar(mapping=aes( x=CARRIER, fill=Delayed),position="dodge") + ylab("Flight Count") + ggtitle("Total Flights vs Delayed Flights by Carrier") + labs(caption  = "*Does not include cancelled or diverted flights*") + theme_minimal()
```

**Findings**

With a probability of .3, B9 (BIMINI Island Air Inc) is the carrier with the most delayed arrival times. This means that 30% of all their flights that land in Denver arrive late. So if your goal is to arrive in Denver on time I recommend flying UA (United Airlines) because they only have a 13% chance of delayed arrival. 

# Description of What Each Member Did 
<p>

## Julia Foley 
<p>
**Individual**
For the individual part, I studied the probability that a flight would have an arrival delay, a departure delay, and a late aircraft delay. I focused on delays that were after the flight was supposed to be rather than delays that left early. I started by calculating the number of delayed flights arriving or departing in Denver compared to the number of flights total arriving or departing in Denver. Then, I found the probability that a flight would have a depature delay and a late aircraft delay given that it has an arrival delay out of the total number of flights arriving or departing in Denver. Next, I found the probability that a flight would have a depature delay and a late aircraft delay given that it has an arrival delay out of the total number of flights with an arrival delay arriving or departing in Denver. Lastly, I found the probability that a flight would have a depature delay and a late aircraft delay given that it has an arrival delay out of the total number of flights in the data set. For the plot, I plotted arrival delay by departure delay. I then used mutate to create a new variable to represent whether or not there is a late aircraft delay. I, then, used fact grid with the new variable to seperate the flights with a late aircraft delay and the flights without a late aircraft delay. 

**Team**

For the team part, our team got together on Thursday and together came up with our two main questions and how we were gonna do the plots. As a team we created a hypothesis for question 2. Then, I began working on question 2. I first calculated the probability that a flight arrived or departed in denver out of the total number of flights. Then, for each time, I found the probability that there would be a delay at that time out of the total number of flights arriving or departing in Denver at that time. Additionally, for each time, I used mutate to create a new variable that would say whether or not there was a arrival delay. I, then, created a bar plot to represent how many flights there were at the specific time and seperated whether or not there was a delay using the new variable. 
<p>

##Nisia Pinto
**Team**

Helped formulate the one of the questions related to best time to travel.

**Individual**

Used probabilities to find which one of the airline companies travelling to and from Denver have more delay. I also plotted the density graphs.

##Andres Acevedo
<p>
**Individual**

**Hypothesis**
Longer flights will have more time in the air to make up any lost time from a departure or a weather delay. therefore, the arrival delays associated with very long distance >2000 miles flights will be less than the arrival delays associated with moderately long distance >1000 mile flights. 

```{r, echo = TRUE}

ggplot(LongDistance)  +
 geom_histogram(fill = "#FF6666", binwidth = 10, mapping =aes(x = ARR_DELAY))+
  xlim(-50,300) +
  ggtitle('Distribution of Long Distance Flights and their Arrival Delays') + 
  xlab('Arrival Delay in Minutes') + 
  ylab('Number of Flights')

#First, find all FLights that were not cancelled and were headed to denver.
DENarrivals <- COflights %>% filter(DEST =='DEN', CANCELLED == 0)

#Now Filter DENarrivals to find the number of flights with ARR_DELAY >15
DENdelays <- DENarrivals %>% filter(ARR_DELAY > 15)

summarize(DENarrivals, 'Number of Flights into DEN' = n())
summarize(DENdelays, 'Number of Delayed Flights into DEN' = n())
#Probability A = 
(33567/221642) *100
#Now count the number of delayed flights into denver and divide it by total flights arriving in denver (Number of Delayed Flights arriving in denver/Number of flights arrived in denver) this will be your probability of a flight into denver being delayed.




#Joint Probability is 

#Number of Flights into DEN w/ARR_Delay >60/Number of Delayed Flights Arriving in Denver
#For your joint probability you want to take DENdelays and filter is to include only ARR_Delay>60
DENBigDelay <- DENdelays %>% filter(ARR_DELAY >60)

summarize(DENBigDelay, 'number of flights delayed 60 min or more' = n())

summarize(DENdelays, 'Number of Delayed Flights arrived in Denver' = n())
#Joint Probability 
(10541/33567) *100


#Conditional Probability. 

#Looking for the probability that a flight into denver will be delayed by more than 60 minutes given that it traveled more than 1000 miles.

#First Define a new data set that includes all flights into denver that are delayed by more than an hour and also came from a distance of more than 1000 miles.
BigDelay <- DENdelays %>% filter(ARR_DELAY >60, DISTANCE >1000)
LongDistance <- DENarrivals %>% filter(DISTANCE >1000)

summarize(BigDelay, 'Num of flights delayed more than 60min' = n())
summarize(LongDistance, 'Num of flights with Distance >1000' = n())
#Condition Probability that a Filight will arrive more than an hour late given that it came more than 1000 miles
(3188/64620)*100 

I explored the relationhsip between the distance traveled by a plane and it's arrival delay by calculating the probability that a flight into Denver would be delayed in arriving. I then filtered the dataset to reflect the arrival delays longer than an hour. Finally, to see what the probability was that a flight travelling 1000 miles or more would be delayed arriving my 60 or more minutes, I calculated conditional probability 

**Team**

I created the plot for Question 1. I used the HolidaySeason data frame to initialize the ggplot. From there, I used a histogram to count the number of delayed flights arriving in Denver in a given month. I filled the bars with the 'steelblue' aesthetic and outlined the bins with the 'turquoise2' aesthetic. I set the bin widths to 1 to represent each month and scaled the x-axis to be between [0,12] with a step of 1 month. I added a title as well as labels and labels as well as a minimal theme. 


## Declan Franklin
<p>

**Individual** 


**Individual** 



## Thomas Neal
<p>

**Individual**


At first, I was looking into what causes flights to be canceled by comparing total canceled flights to the cancellation code. However, although I found interesting data I decided not to focus on this because It didn't help progress our team's research. I then pivoted to look at the probability that a carrier would have a canceled flight. This also provided interesting data, however, I decided to change my research question one last time to the probability that a carrier will have a delayed arrival time. This allowed me to make recommendations based on the carrier, and better related to our team question.

#### Non-Defaults

A custom data frame, fill, postion=" dodge", title, x and y labels, caption, theme.

**Team**

I helped to alter the data we were looking at in order to better represent the questions we were researching.


