---
title: "Lab 8"
author: "P-Hacker: Andres Acevedo, Julia Foley, Thomas Neal, Nisia Pinto, Declan Franklin"
date: "10/17/2019"
output: html_document
---

## Team
In this lab we will find analyze ...

### Domain Expert:
Our domain expert is...FIFA, sports recruiters, football betting...
### Overall Question
Question is...

### Why Question is Important

### Why Those Reasons are Important

### Dataset Background
The dataset we will analyze next week is data for the 2015-2016 season of the English Premier League. This data includes statistics relating to soccer such as Home Goals for each game and Number of Red Cards for each game. 

## Individual Questions

### Andres Acevedo:
**Questions:**
Why does telegraph.co hate these refs so much? Do they really do aything wrong? Or are their mistakes just overpublicized? I'll try to answer these questions by looking at the number of penalties called by the 'Best' and 'Worst' refs in the Barclays Premier League. I want to see if the bad refs call a disproportinate number of penalties compared to the good refs. I want to see if they give out more yellow cards or red cards. If the home team wins too much or too little. If there a greater number of goals scored or if fewer goals are scored. My backup question is to find out how penalties impact the number of goals scored in a game. If a higher number of penalties resulted in more or less total goals being scored. 

**Why are These Questions Interesting/Important?**
Question 1: Im interested in this question because it's always good to have a way to gague the veracity of a news article.  If these refs are considered to be the 'worst' or 'best' refs, teams need to know if theres any reason for that. We as readers should know whether to take the premise of the article seriously. 

Question2:
There are many kinds of penalties. Some will result in a direct scoring chance like from a cornerkick or a free kick near the goalie box. Others will result in a change of posession which could have little chance of resulting in goal. Some even result in an ejection. Teams have to make informed decisions about how aggressively to play their opponents. If the number of penalties doesnt result in fewer goals scored by the team, or more goals scored by the opposing team, it may be worth it to play more aggressively. If the opposite, it may be reasonable to back off and just crowd the box with defense to take away passing lanes.


### Julia Foley:
**Questions:**
Question 1: Which team has the highest probability of winning when they are at home?
Question 2: Which referee gave the most red cards? 

**Why are These Questions Interesting/Important?**
Question 1: This is interesting because it will show whether or not a team will win at home. I think that it is interesting to see if there is a trend between winnings and location. It is interesting to see if being the home team impacts the probability of a win. Additionally, it will be interesting to see if a team with a better record overall has a higher probability of winning at home or if there is another factor that goes into it. This  question is important because it could help to explain the idea of home team advantage. Additionally, it could help determine the win average and therefore, help predict whether or not a team is going win. This could help teams determine whether or not the home team advantage plays a role in their win average. Lastly, it will show which teams are better when they are at home. 
Question 2: I find it interesting to see which referees gave out more red cards because it shows which referees are stricter. It is interesting to see if the number of red cards is correlated by the number of games each referee refereed or if there is an outside factor that contributes to the correlation. I think it is fascinating to see the data and reasons behind the number of red cards. This is important because it shows if a certain referee is more likely to give a player a red card. If this is the case, the team may want to be more careful about fouls to avoid red cards. Additionally, it can show how strict a referee is which would show whether a referee stays to the rules exactly or whether he is more lenient. This could change how some of the games are played if the players know how strict the referee is. 

### Thomas Neal:
**Questions:**
Question1:
Question2:

**Why are These Questions Interesting/Important?**
Question1:
Question2:

### Declan Franklin:

**Questions:**
Question1:
Question2:

**Why are These Questions Interesting/Important?**
Question1:
Question2:

### Nisia Pinto:
**Questions:**
Question1:
Question2:

**Why are These Questions Interesting/Important?**
Question1:
Question2:

## Visualizations of the East Africa Lake Data

### Andres Acevedo:

**Tanganyka (Done in class with Dr.Vance)**
```{r}
library(tidyverse)
library(readxl)
path <- "/Users/Andres/Desktop/DataScience2600/East Africa lake data.xlsx"
Tang <- read_excel(path, sheet = 6)

Tang <- select(Tang, `Year AD...1`,`BSi (normalized)`,`Year AD...4`,`Charcoal (normalized)`,`Year`,TEX86)

Tang <- rename(Tang,year1=`Year AD...1`,BSi=`BSi (normalized)`, year2=`Year AD...4`,Charcoal=`Charcoal (normalized)`,year3=`Year`)

Tang1 <- Tang %>% select(year1,BSi)
Tang2 <- Tang %>% select(year2,Charcoal)
Tang3 <- Tang %>% select(year3,TEX86)

Tang1$year1 <- round(Tang1$year1)
Tang2$year2 <- round(Tang2$year2)
Tang3$year3 <- round(Tang3$year3)

Tang12 <- full_join(Tang1,Tang2,by=c("year1"="year2"))

Tang <- full_join(Tang12,Tang3,by = c("year1"="year3"))

Tang.skinny <- gather(Tang,key="measurement",value="value",BSi, Charcoal,TEX86)

ggplot(Tang.skinny) +
  geom_line(mapping = aes(x = year1, y = value), color = 'steelblue') + facet_wrap(~measurement) +
  ggtitle("Rainfall Variation vs Time for Tang") + xlab('Year') + ylab('Rainfall Variation') + theme_minimal() + scale_x_continuous(breaks = seq(1000, 2000, 250)) + scale_y_continuous(breaks = seq(0, 2.5, 0.5))

ggplot(Tang.skinny) +
  geom_smooth(mapping = aes(x = year1, y = value), color = 'black') + geom_jitter(mapping = aes(x = year1, y = value), color = 'steelblue', alpha = 0.5) + facet_wrap(~measurement) + ggtitle('Rain Variation Vs Time for Tang') + xlab('Year') + ylab('Rainfall Variation')
```

### Julia Foley
**VOI Rainfall**

```{r}
VOI <- select(voi2,`Year`, `Jan`, `Feb`, `Mar`, `Apr`, `May`, `Jun`, `Jul`, `Aug`, `Sep`, `Oct`, `Nov`, `Dec`)
VOI <- gather(VOI,Jan,Feb,Mar,Apr,May,Jun,Jul,Aug,Sep,Oct,Nov,Dec,key=month,value=rainfall)

ggplot(VOI, mapping = aes(Year,rainfall)) +
  geom_count() +
  labs(y = "Amount of Rainfall in a Month")
```

**Thomas Neal**
**Challa Rainfall**

```{r, echo=FALSE}
library(readxl)
Challa <- East_Africa_lake_data <- read_excel("East Africa lake data.xlsx", sheet = "Challa")
Challa=Challa[-1,]

Challa[is.na(Challa)] <- 0

Challa <- Challa %>% mutate( `varve thickness`=as.double(`varve thickness`),`dark layer`=as.double(`dark layer`), `light layer`=as.double(`light layer`) )



 ggplot(data=Challa) + geom_smooth(mapping=aes(x=`varve number`, y=`dark layer`, color="dark layer"), se=FALSE)  + geom_smooth(mapping=aes(x=`varve number`, y=`varve thickness`, color="varve thickness" ),se=FALSE) + geom_smooth(mapping=aes(x=`varve number`, y=`light layer`,color="light layer" ),se=FALSE) + labs(title="Rainfall in Challa over time", x="Time", y="amount of rain") +  scale_colour_manual("", 
                      breaks = c("dark layer", "varve thickness", "light layer"),
                      values = c("red", "green", "blue"))+ theme_dark()


```

### Declan Franklin
**xxx**

### Nisia Pinto
**CRU and MAM Rainfall Data**
